### 2015-02-19 doc/proposal/proposal.tex
Initial draft for proposal

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amssymb}

\usepackage{hyperref}  % for urls and hyperlinks

\setlength{\textwidth}{6.2in}
\setlength{\oddsidemargin}{0.3in}
\setlength{\evensidemargin}{0in}
\setlength{\textheight}{8.7in}
\setlength{\voffset}{-.7in}
\setlength{\headsep}{26pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

% input some useful macros from RJLmacros.tex:
\input{../macros/RJLmacros}

\begin{document}
\section{Description}
We are planing to model the interaction of a bore and a free-standing coastal structure with Clawpack in 2D. In a first step we plan to model the wave tank followed by placing the coastal structure, a square column in the domain. We would like to compare the results with experimental results by Halld\'or \'Arnason.

\end{document}

### 2015-02-20 doc/proposal/proposal.tex
modified proposal

@@
 We are planing to model the interaction of a bore and a free-standing coastal structure with Clawpack in 2D.
-In a first step we plan to model the wave tank followed by placing the coastal structure, a square column in the domain.
-We would like to compare the results with experimental results by Halld\'or \'Arnason.
+Simulation work can be compared to benchmark experiment by Halld\'or \'Arnason.
+The benchmark experiment was conducted with a 16.6 m long, 0.6m wide and 0.45 m deep wave tank. When a gate is lifted, it generates a single bore to impinge the column located downstream.
+We are going to simulate 2 cases in the experiment. The 1st case purely generated the single bore, the 2nd case generated the bore to impinge the column with one side facing the flow.
+\par
+In 1st case, wave height and streamwise velocity, U, at different depth, at where the center of column should located should be measured.
+In the 2nd, streamwise forces on the column and wave height at some place in front of the column should be measured.
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=0.5\textwidth]{./figures/Diagram_case}
+            \caption{Diagram of experiment apparatus}
+        \end{figure}
 
+\section{Implementation steps}
+In a first step we plan to study some relavant chapters in textbook, including multidimensional hyperbolic problems, multidimensional numerical methods and implementation of multidimensional system in CLAWPACK, especially the shallow water equations.
+\par
+Next we plan to model a dambreak problem in CLAWPACK, using shallow water equation.
+\par
+Then we will investigate on how to put a square column into the mesh in CLAWPACK.
+Maybe cutting out a section of the grid where the column is and modify the code to make it treat the interface as boundary.
+\par
+Finally all simulation result will be compared to experiment result.
+\section{validation procedure}
+The validation procedure will include comparison of wave height and streamwise velocity, U, at different depth, at where the center of column should located in case without a column, and comparison of wave height in front of the column.
 \end{document}

### 2015-02-20 doc/proposal/proposal.tex
Final proposal

@@
 \begin{document}
+\section{Introduction}
+We are planing to model the interaction of a bore and a free-standing coastal structure with CLAWPACK in 2D.
+Simulation work can be compared to benchmark experiment by Halld\'or \'Arnason \cite{HA}.
 \section{Description}
-We are planing to model the interaction of a bore and a free-standing coastal structure with Clawpack in 2D.
-Simulation work can be compared to benchmark experiment by Halld\'or \'Arnason.
-The benchmark experiment was conducted with a 16.6 m long, 0.6m wide and 0.45 m deep wave tank. When a gate is lifted, it generates a single bore to impinge the column located downstream.
+The benchmark experiment was conducted in a 16.6 m long, 0.6m wide and 0.45 m deep wave tank. When a gate is lifted, it generates a single bore to impinge the column located downstream.
 We are going to simulate 2 cases in the experiment. The 1st case purely generated the single bore, the 2nd case generated the bore to impinge the column with one side facing the flow.
 \par
-In 1st case, wave height and streamwise velocity, U, at different depth, at where the center of column should located should be measured.
-In the 2nd, streamwise forces on the column and wave height at some place in front of the column should be measured.
+In the 1st case, wave height and streamwise velocity, U, at different depth, at where the center of column should be located are measured.
+In the 2nd case, the wave height in front and behind the column are measured.
         \begin{figure}[h!]
@@
 \par
-Next we plan to model a dambreak problem in CLAWPACK, using shallow water equation.
+Next we plan to model a dambreak problem in CLAWPACK, using the shallow water equation.
 \par
@@
 Finally all simulation result will be compared to experiment result.
-\section{validation procedure}
-The validation procedure will include comparison of wave height and streamwise velocity, U, at different depth, at where the center of column should located in case without a column, and comparison of wave height in front of the column.
+\section{Validation Procedure}
+The validation procedure will include comparison of wave height and streamwise velocity, U, at different depth, at where the center of column should be located in case without a column, and comparison of wave height in front of the column.
+\end{itemize}
+{\footnotesize
+\begin{thebibliography}{100}
+\bibitem{HA} Halld\'or \'Arnason  (2005). Thesis - Interactions between an incident bore and a free-standing coastal structure. \url{http://faculty.washington.edu/cpetroff/Halldor%20dissertation.pdf}
+\end{thebibliography}
+}
 \end{document}

### 2015-03-02 dambreak/maketopo.py
set topo to zero

@@ -48,9 +48,8 @@ def topo(x,y):
     zz = where(zz<hi, zz, hi)
     z = where(dist<base_r, zz, 0.)
 
+    #z = z - 0.054
     z = zeros(x.shape)
-
-#    z = z - 0.054
     
     return z
 

### 2015-03-02 dambreak/qinit.f90
qinit.f set to correct water tank level

@@ -13,7 +13,7 @@ subroutine qinit(meqn,mbc,mx,my,xlower,ylower,dx,dy,q,maux,aux)
     real(kind=8), intent(inout) :: aux(maux,1-mbc:mx+mbc,1-mbc:my+mbc)
 
     !try to define sloc,hl,ul,hr,ur locally
-    REAL(kind=8) :: sloc = 1.9
+    REAL(kind=8) :: sloc = 5.9
     REAL(kind=8) :: hl = 0.25 
     REAL(kind=8) :: ul = 0.0
     REAL(kind=8) :: hr = 0.02

### 2015-03-02 dambreak/setrun.py
added guages for column center

@@ -32,24 +32,6 @@ def setrun(claw_pkg='geoclaw'):
     num_dim = 2
     rundata = data.ClawRunData(claw_pkg, num_dim)
 
-
-
-    #------------------------------------------------------------------
-    # Problem-specific parameters to be written to setprob.data:
-    #------------------------------------------------------------------
-    #For dambreak I.C
-    probdata = rundata.new_UserData(name='probdata',fname='setprob.data')
-
-    #probdata.add_param('g',   1.0,  'Gravitational constant')
-    probdata.add_param('sloc',    5.9,   'Initial discontinuity location')
-    probdata.add_param('hl', 0.25,  'Value of water height to left')
-    probdata.add_param('ul', 0.,  'Value of water velocity to right')
-    probdata.add_param('hr', 0.02,  'Value of water height to left')
-    probdata.add_param('ur', 0.,  'Value of water velocity to right')
-
-
-
-
     #------------------------------------------------------------------
     # GeoClaw specific parameters:
     #------------------------------------------------------------------
@@ -80,11 +62,17 @@ def setrun(claw_pkg='geoclaw'):
     clawdata.lower[1] = 0
     clawdata.upper[1] = 0.6
 
+    #clawdata.lower[0] = 0.0
+    #clawdata.upper[0] = 19.76  #9.84
+
+    #clawdata.lower[1] = 0.
+    #clawdata.upper[1] = 1.52
+
 
 
     # Number of grid cells: Coarsest grid
-    clawdata.num_cells[0] = 332 
-    clawdata.num_cells[1] = 12 
+    clawdata.num_cells[0] = 260
+    clawdata.num_cells[1] = 12
 
 
     # ---------------
@@ -131,8 +119,8 @@ def setrun(claw_pkg='geoclaw'):
 
     if clawdata.output_style==1:
         # Output nout frames at equally spaced times up to tfinal:
-        clawdata.num_output_times =10 
-        clawdata.tfinal = 2.
+        clawdata.num_output_times = 40
+        clawdata.tfinal = 12.0 
         clawdata.output_t0 = True  # output at initial (or restart) time?
 
     elif clawdata.output_style == 2:
@@ -161,7 +149,7 @@ def setrun(claw_pkg='geoclaw'):
[…]

### 2015-03-02 dambreak_new/b4step2.f90
delete dambreak_new folder

@@ -1,72 +1,0 @@
-! ============================================
-subroutine b4step2(mbc,mx,my,meqn,q,xlower,ylower,dx,dy,t,dt,maux,aux)
-! ============================================
-!
-! # called before each call to step
-! # use to set time-dependent aux arrays or perform other tasks.
-!
-! This particular routine sets negative values of q(1,i,j) to zero,
-! as well as the corresponding q(m,i,j) for m=1,meqn.
-! This is for problems where q(1,i,j) is a depth.
-! This should occur only because of rounding error.
-!
-! Also calls movetopo if topography might be moving.
-
-    use geoclaw_module, only: dry_tolerance
-    use geoclaw_module, only: g => grav
-    use topo_module, only: num_dtopo,topotime
-    use topo_module, only: aux_finalized
-    use topo_module, only: xlowdtopo,xhidtopo,ylowdtopo,yhidtopo
-
-    use amr_module, only: xlowdomain => xlower
-    use amr_module, only: ylowdomain => ylower
-    use amr_module, only: xhidomain => xupper
-    use amr_module, only: yhidomain => yupper
-    use amr_module, only: xperdom,yperdom,spheredom,NEEDS_TO_BE_SET
-
-    implicit none
-
-    ! Subroutine arguments
-    integer, intent(in) :: meqn
-    integer, intent(inout) :: mbc,mx,my,maux
-    real(kind=8), intent(inout) :: xlower, ylower, dx, dy, t, dt
-    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc,1-mbc:my+mbc)
-    real(kind=8), intent(inout) :: aux(maux,1-mbc:mx+mbc,1-mbc:my+mbc)
-
-    ! Local storage
-    integer :: index,i,j,k,dummy
-    real(kind=8) :: h,u,v,x
-
-    ! Check for NaNs in the solution
-    call check4nans(meqn,mbc,mx,my,q,t,1)
-
-    ! check for h < 0 and reset to zero
-    ! check for h < drytolerance
-    ! set hu = hv = 0 in all these cells
-    forall(i=1-mbc:mx+mbc, j=1-mbc:my+mbc,q(1,i,j) < dry_tolerance)
-        q(1,i,j) = max(q(1,i,j),0.d0)
-        q(2:3,i,j) = 0.d0
-    end forall
-
-
-    if (aux_finalized < 2) then
-        ! topo arrays might have been updated by dtopo more recently than
[…]

### 2015-03-02 dambreak_new/bc2amr.f

@@ -1,327 +1,0 @@
-
-c
-c ------------------------------------------------------------------
-c
-      subroutine bc2amr(val,aux,nrow,ncol,meqn,naux,
-     1                  hx, hy, level, time,
-     2                  xleft,  xright,  ybot, ytop,
-     3                  xlower, ylower,xupper,yupper,
-     4                  xperiodic, yperiodic,spheredom)
-
-c
-c    Specific to geoclaw:  extrapolates aux(i,j,1) at boundaries
-c    to constant.
-c
-c :::::::::: bc2amr ::::::::::::::::::::::::::::::::::::::::::::::;
-c
-c     Take a grid patch with mesh widths hx,hy, of dimensions nrow by
-c     ncol,  and set the values of any piece of
-c     of the patch which extends outside the physical domain
-c     using the boundary conditions.
-c
-c     ------------------------------------------------
-c     # Standard boundary condition choices for amr2ez in clawpack
-c
-c     # At each boundary  k = 1 (left),  2 (right),  3 (top), 4 (bottom):
-c     #   mthbc(k) =  0  for user-supplied BC's (must be inserted!)
-c     #            =  1  for zero-order extrapolation
-c     #            =  2  for periodic boundary coniditions
-c     #            =  3  for solid walls, assuming this can be implemented
-c     #                  by reflecting the data about the boundary and then
-c     #                  negating the 2'nd (for k=1,2) or 3'rd (for k=3,4)
-c     #                  component of q.
-c     #            =  4  sphere bcs (left half maps to right half of same
-c     #                  side, and vice versa), as if domain folded in half
-c     ------------------------------------------------
-c
-c     The corners of the grid patch are at
-c        (xleft,ybot)  --  lower left corner
-c        (xright,ytop) --  upper right corner
-c
-c     The physical domain itself is a rectangle bounded by
-c        (xlower,ylower)  -- lower left corner
-c        (xupper,yupper)  -- upper right corner
-c
-c     the picture is the following:
-c
-c               _____________________ (xupper,yupper)
-c              |                     |
-c          _________ (xright,ytop)   |
-c          |   |    |                |
-c          |   |    |                |
-c          |   |    |                |
[…]

### 2015-03-02 dambreak_new/compare_gauges.py

@@ -1,77 +1,0 @@
-
-
-from pylab import *
-from clawpack.visclaw.data import ClawPlotData
-
-plotdata = ClawPlotData()
-
-if 0:
-    plotdata.outdir = '_output'
-    toffset = 0.
-
-if 0:
-    plotdata.outdir = '_output_manning_0.025'
-    toffset = 92.
-if 0:
-    plotdata.outdir = '_output_manning015_cfl090'
-    toffset = 92.
-if 1:
-    plotdata.outdir = '_output_manning015_cfl089'
-    toffset = 96.
-
-figure(50,figsize=(18,12))
-clf()
-
-# ---  Gauge 1 ---
-
-d = loadtxt('s1u.txt')
-t1u = d[:,0]
-s1u = d[:,1]
-
-d = loadtxt('s1v.txt')
-t1v = d[:,0]
-s1v = d[:,1]
-
-g = plotdata.getgauge(1)
-u = g.q[1,:] / g.q[0,:]
-v = g.q[2,:] / g.q[0,:]
-
-
-subplot(4,1,1)
-plot(t1u+toffset,s1u,'b',label='Experiment')
-plot(g.t, u, 'r',label='GeoClaw')
-ylabel('u (m/s)')
-legend(loc='upper right')
-
-subplot(4,1,2)
-plot(t1v+toffset,s1v,'b',label='Experiment')
-plot(g.t, v, 'r',label='GeoClaw')
-ylabel('v (m/s)')
-
-# ---  Gauge 2 ---
-
-d = loadtxt('s2u.txt')
-t2u = d[:,0]
-s2u = d[:,1]
-
-d = loadtxt('s2v.txt')
-t2v = d[:,0]
-s2v = d[:,1]
-
-g = plotdata.getgauge(2)
-u = g.q[1,:] / g.q[0,:]
-v = g.q[2,:] / g.q[0,:]
-
-subplot(4,1,3)
-plot(t2u+toffset,s2u,'b',label='Experiment')
-plot(g.t, u, 'r',label='GeoClaw')
-ylabel('u (m/s)')
-legend(loc='upper right')
-
-subplot(4,1,4)
-plot(t2v+toffset,s2v,'b',label='Experiment')
-plot(g.t, v, 'r',label='GeoClaw')
-ylabel('v (m/s)')
-
-show()
-#savefig('gauges_manning015_cfl089.png')

### 2015-03-02 dambreak_new/make_BM1.py

@@ -1,88 +1,0 @@
-"""
-Create the BM1_leveque.txt file requested by Pat Lynett.
-"""
-
-from pylab import *
-from clawpack.visclaw.data import ClawPlotData
-
-plotdata = ClawPlotData()
-
-toffset = 0.
-
-if 1:
-    plotdata.outdir = '_output_manning010_cfl090'
-    fname = 'BM1_leveque_1.txt'
-if 0:
-    plotdata.outdir = '_output_manning010_cfl089'
-    fname = 'BM1_leveque_2.txt'
-if 0:
-    plotdata.outdir = '_output_manning015_cfl090'
-    fname = 'BM1_leveque_3.txt'
-if 0:
-    plotdata.outdir = '_output_manning015_cfl089'
-    fname = 'BM1_leveque_4.txt'
-
-figure(figsize=(8,12))
-clf()
-
-# ---  Gauge 1 ---
-
-d = loadtxt('s1u.txt')
-t1u = d[:,0]
-s1u = d[:,1]
-
-d = loadtxt('s1v.txt')
-t1v = d[:,0]
-s1v = d[:,1]
-
-g = plotdata.getgauge(1)
-u1 = g.q[1,:] / g.q[0,:]
-v1 = g.q[2,:] / g.q[0,:]
-t1 = g.t
-
-subplot(4,1,1)
-plot(t1u+toffset,s1u,'b',label='Experiment')
-plot(t1, u1, 'r',label='GeoClaw')
-ylabel('u (m/s)')
-legend(loc='upper right')
-title(plotdata.outdir)
-
-subplot(4,1,2)
-plot(t1v+toffset,s1v,'b',label='Experiment')
-plot(t1, v1, 'r',label='GeoClaw')
-ylabel('v (m/s)')
-
-# ---  Gauge 2 ---
-
-d = loadtxt('s2u.txt')
-t2u = d[:,0]
-s2u = d[:,1]
-
-d = loadtxt('s2v.txt')
-t2v = d[:,0]
-s2v = d[:,1]
-
-g = plotdata.getgauge(2)
-u2 = g.q[1,:] / g.q[0,:]
-v2 = g.q[2,:] / g.q[0,:]
-t2 = g.t
-
-
-
-subplot(4,1,3)
-plot(t2u+toffset,s2u,'b',label='Experiment')
-plot(t2, u2, 'r',label='GeoClaw')
-ylabel('u (m/s)')
-legend(loc='upper right')
[…]

### 2015-03-02 dambreak_new/maketopo.py

@@ -1,72 +1,0 @@
-
-from clawpack.geoclaw import topotools
-from pylab import *
-
-dx=.0025;
-
-def maketopo():
-    """
-    Output topography files
-    """
-    nxpoints = 201
-    nypoints = 21
-    xlower = 0.
-    xupper = 20.
-    ylower = 0.
-    yupper = 1.52
-    outfile= "domain.tt1"
-    topotools.topo1writer(outfile,topo,xlower,xupper,ylower,yupper,nxpoints,nypoints)
-
-    nxpoints = 201
-    nypoints = 201
-    xlower = 4.5
-    xupper = 5.5
-    ylower = 0.25
-    yupper = 1.25
-    outfile= "hump.tt1"
-    topotools.topo1writer(outfile,topo,xlower,xupper,ylower,yupper,nxpoints,nypoints)
-
-
-def topo(x,y):
-    """
-    flat with hump
-    """
-
-    z = zeros(x.shape)
-
-    # island addition
-    # from paper
-    xo=5.0;
-    yo=1.52/2;
-    slope_d=8;
-    base_r=0.75/2;
-    top_r=0.05/2;
-    hi=0.049;
-
-    dist = sqrt( (x-xo)**2+(y-yo)**2 );
-    zz = hi*(1-(dist-top_r)/(base_r-top_r))
-    zz = where(zz<hi, zz, hi)
-    z = where(dist<base_r, zz, 0.)
-
-    z = z - 0.054
-    
-    return z
-
-def plot_topo():
-
-    topo1 = topotools.Topography()
-    topo1.read('domain.tt1',1)
-    figure(figsize=(12,4))
-    ax = axes()
-    topo1.plot(axes=ax,add_colorbar=False)
-
-    topo2 = topotools.Topography()
-    topo2.read('hump.tt1',1)
-    topo2.plot(axes=ax)
-    ax.set_xlim(0,9.84)
-    ax.set_ylim(0,1.52)
-    axis('scaled')
-
-if __name__=='__main__':
-    maketopo()
-    

### 2015-03-02 dambreak_new/qinit.f90

@@ -1,61 +1,0 @@
-
-subroutine qinit(meqn,mbc,mx,my,xlower,ylower,dx,dy,q,maux,aux)
-    
-    use qinit_module, only: qinit_type,add_perturbation
-    use geoclaw_module, only: sea_level
-    
-    implicit none
-    
-    ! Subroutine arguments
-    integer, intent(in) :: meqn,mbc,mx,my,maux
-    real(kind=8), intent(in) :: xlower,ylower,dx,dy
-    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc,1-mbc:my+mbc)
-    real(kind=8), intent(inout) :: aux(maux,1-mbc:mx+mbc,1-mbc:my+mbc)
-
-    !try to define sloc,hl,ul,hr,ur locally
-    REAL(kind=8) :: sloc = 5.9
-    REAL(kind=8) :: hl = 0.25 
-    REAL(kind=8) :: ul = 0.0
-    REAL(kind=8) :: hr = 0.02
-    REAL(kind=8) :: ur = 0.0
-    REAL(kind=8) :: xcenter
-
-    !common /comic/ sloc,hl,ul,hr,ur
-    
-    ! Locals
-    integer :: i,j,m
-    real(kind=8) :: discharge
-
-    !print *,"----------------executing subroutine qinit()------------------------"
-    
-    ! Set flat state based on sea_level
-    q = 0.d0
-    !discharge = 0.115d0 * 0.054  ! free stream
-    do i=1,mx
-        xcenter = xlower + (i - 0.5d0) * dx
-        do j=1,my
-            if (xcenter < sloc) then
-                q(1,i,j) = hl
-                q(2,i,j) = hl*ul
-            else
-                q(1,i,j) = hr
-                q(2,i,j) = hr*ur
-            endif
-        enddo
-    enddo
-
-!The grammar below seem to be incorrect for *.f90. It's only correct fo *.f
-!    forall(i=1:mx, j=1:my)
-!        xcenter = xlower + (i - 0.5d0) * dx
-!        if (xcenter < sloc) then
-!            q(1,i,j) = hl
-!            q(2,i,j) = hl*ul
-!        else
-!            q(1,i,j) = hr
-!            q(2,i,j) = hr*ur
-!        !q(1,i,j) = max(0.d0, sea_level - aux(1,i,j))
-!        !q(2,i,j) = discharge   ! 0.115d0 * q(1,i,j)
-!        endif
-!    end forall
-
-end subroutine qinit

### 2015-03-02 dambreak_new/save_data.m

@@ -1,28 +1,0 @@
-
-% convert data to ascii text files for reading into Python:
-
-load ../all_data/1_Submerged_Island_Lab/comparison_data/SL_S1_U.dat;
-te1 = SL_S1_U(:,1);
-ue1 = SL_S1_U(:,2);
-s1u = [te1, ue1];
-save('S1u.txt','s1u','-ascii')
-
-load ../all_data/1_Submerged_Island_Lab/comparison_data/SL_S1_V.dat;
-te2 = SL_S1_V(:,1);
-ve2 = SL_S1_V(:,2);
-s1v = [te2, ve2];
-save('S1v.txt','s1v','-ascii')
-
-load ../all_data/1_Submerged_Island_Lab/comparison_data/SL_S2_U.dat;
-te3 = SL_S2_U(:,1);
-ue3 = SL_S2_U(:,2);
-s2u = [te3, ue3];
-save('S2u.txt','s2u','-ascii')
-
-load ../all_data/1_Submerged_Island_Lab/comparison_data/SL_S2_V.dat;
-te4 = SL_S2_V(:,1);
-ve4 = SL_S2_V(:,2);
-s2v = [te4, ve4];
-save('S2v.txt','s2v','-ascii')
-
-

### 2015-03-02 dambreak_new/setplot.py

@@ -1,502 +1,0 @@
-
-""" 
-Set up the plot figures, axes, and items to be done for each frame.
-
-This module is imported by the plotting routines and then the
-function setplot is called to set the plot parameters.
-    
-""" 
-
-import numpy
-from clawpack.visclaw import colormaps
-
-
-
-#--------------------------
-def setplot(plotdata):
-#--------------------------
-    
-    """ 
-    Specify what is to be plotted at each frame.
-    Input:  plotdata, an instance of pyclaw.plotters.data.ClawPlotData.
-    Output: a modified version of plotdata.
-    
-    """ 
-
-
-    from clawpack.visclaw import colormaps, geoplot
-
-    plotdata.clearfigures()  # clear any old figures,axes,items data
-    plotdata.format = 'binary'
-
-    def set_drytol(current_data):
-        # The drytol parameter is used in masking land and water and
-        # affects what color map is used for cells with small water depth h.
-        # The cell will be plotted as dry if h < drytol.
-        # The best value to use often depends on the application and can
-        # be set here (measured in meters):
-        current_data.user['drytol'] = 1.e-2
-
-    plotdata.beforeframe = set_drytol
-
-    # To plot gauge locations on imshow or contour plot, use this as
-    # an afteraxis function:
-
-    def addgauges(current_data):
-        from clawpack.visclaw import gaugetools
-        gaugetools.plot_gauge_locations(current_data.plotdata, \
-             gaugenos='all', format_string='ko', add_labels=True, fontsize=8)
-    
-    #-----------------------------------------
-    # Figure for imshow plot
-    #-----------------------------------------
-    plotfigure = plotdata.new_plotfigure(name='surface', figno=0)
-    plotfigure.kwargs = {'figsize':(14,10)}
-
-    # Set up for axes in this figure:
-    plotaxes = plotfigure.new_plotaxes('surface')
-    plotaxes.axescmd = 'subplot(311)'
-    #plotaxes.axescmd = 'axes([.1,.7,.8,.2])'
-    plotaxes.title = 'Surface'
[…]

### 2015-03-02 dambreak_new/setplot_compare.py

@@ -1,259 +1,0 @@
-
-""" 
-Set up the plot figures, axes, and items to be done for each frame.
-
-This module is imported by the plotting routines and then the
-function setplot is called to set the plot parameters.
-    
-""" 
-
-import numpy
-from clawpack.visclaw import colormaps
-import os
-
-outdir1 = os.path.abspath('_output_manning010_gt4_cfl089')
-outdir2 = os.path.abspath('_output_manning010_gt4_cfl090')
-
-
-#--------------------------
-def setplot(plotdata):
-#--------------------------
-    
-    """ 
-    Specify what is to be plotted at each frame.
-    Input:  plotdata, an instance of pyclaw.plotters.data.ClawPlotData.
-    Output: a modified version of plotdata.
-    
-    """ 
-
-
-    from clawpack.visclaw import colormaps, geoplot
-
-    plotdata.clearfigures()  # clear any old figures,axes,items data
-    plotdata.format = 'binary'
-    plotdata.outdir = outdir1  # for reading times
-
-    def set_drytol(current_data):
-        # The drytol parameter is used in masking land and water and
-        # affects what color map is used for cells with small water depth h.
-        # The cell will be plotted as dry if h < drytol.
-        # The best value to use often depends on the application and can
-        # be set here (measured in meters):
-        current_data.user['drytol'] = 1.e-2
-
-    plotdata.beforeframe = set_drytol
-
-    # To plot gauge locations on imshow or contour plot, use this as
-    # an afteraxis function:
-
-    def addgauges(current_data):
-        from clawpack.visclaw import gaugetools
-        gaugetools.plot_gauge_locations(current_data.plotdata, \
-             gaugenos='all', format_string='ko', add_labels=True, fontsize=8)
-
-    #-----------------------------------------
-    # Figure for vorticity plot
-    #-----------------------------------------
-
-    plotfigure = plotdata.new_plotfigure(name='vorticity', figno=9)
-    #plotfigure.show = False
-    #plotfigure.kwargs = {'figsize':(16,6)}
-    plotfigure.kwargs = {'figsize':(16,12)}
-
-
[…]

### 2015-03-02 dambreak_new/setrun.py

@@ -1,414 +1,0 @@
-"""
-Module to set up run time parameters for Clawpack.
-
-The values set in the function setrun are then written out to data files
-that will be read in by the Fortran code.
-
-"""
-
-import os
-import numpy as np
-
-
-#------------------------------
-def setrun(claw_pkg='geoclaw'):
-#------------------------------
-
-    """
-    Define the parameters used for running Clawpack.
-
-    INPUT:
-        claw_pkg expected to be "geoclaw" for this setrun.
-
-    OUTPUT:
-        rundata - object of class ClawRunData
-
-    """
-
-    from clawpack.clawutil import data
-
-    assert claw_pkg.lower() == 'geoclaw',  "Expected claw_pkg = 'geoclaw'"
-
-    num_dim = 2
-    rundata = data.ClawRunData(claw_pkg, num_dim)
-
-    #------------------------------------------------------------------
-    # GeoClaw specific parameters:
-    #------------------------------------------------------------------
-    rundata = setgeo(rundata)
-
-    #------------------------------------------------------------------
-    # Standard Clawpack parameters to be written to claw.data:
-    #   (or to amr2ez.data for AMR)
-    #------------------------------------------------------------------
-    clawdata = rundata.clawdata  # initialized when rundata instantiated
-
-
-    # Set single grid parameters first.
-    # See below for AMR parameters.
-
-
-    # ---------------
-    # Spatial domain:
-    # ---------------
-
-    # Number of space dimensions:
-    clawdata.num_dim = num_dim
-
-    # Lower and upper edge of computational domain:
-    clawdata.lower[0] = 0.0
-    clawdata.upper[0] = 16.6
-
-    clawdata.lower[1] = 0
-    clawdata.upper[1] = 0.6
-
-    #clawdata.lower[0] = 0.0
-    #clawdata.upper[0] = 19.76  #9.84
-
-    #clawdata.lower[1] = 0.
-    #clawdata.upper[1] = 1.52
-
-
-
-    # Number of grid cells: Coarsest grid
-    clawdata.num_cells[0] = 260
-    clawdata.num_cells[1] = 12
-
-
-    # ---------------
-    # Size of system:
-    # ---------------
-
-    # Number of equations in the system:
-    clawdata.num_eqn = 3
-
-    # Number of auxiliary variables in the aux array (initialized in setaux)
-    clawdata.num_aux = 1
-
-    # Index of aux array corresponding to capacity function, if there is one:
[…]

### 2015-03-02 dambreak_new/src2.f90

@@ -1,91 +1,0 @@
-subroutine src2(meqn,mbc,mx,my,xlower,ylower,dx,dy,q,maux,aux,t,dt)
-      
-    use geoclaw_module, only: g => grav, coriolis_forcing, coriolis
-    use geoclaw_module, only: friction_forcing, friction_depth
-    use geoclaw_module, only: manning_coefficient
-    use geoclaw_module, only: manning_break, num_manning
-
-    implicit none
-    
-    ! Input parameters
-    integer, intent(in) :: meqn,mbc,mx,my,maux
-    double precision, intent(in) :: xlower,ylower,dx,dy,t,dt
-    
-    ! Output
-    double precision, intent(inout) :: q(meqn,1-mbc:mx+mbc,1-mbc:my+mbc)
-    double precision, intent(inout) :: aux(maux,1-mbc:mx+mbc,1-mbc:my+mbc)
-
-    ! Locals
-    integer :: i, j, nman
-    real(kind=8) :: h, hu, hv, gamma, dgamma, y, fdt, a(2,2), coeff
-    real(kind=8) :: x, q2, q2inflow, xforce
-
-    ! Algorithm parameters
-    ! Parameter controls when to zero out the momentum at a depth in the
-    ! friction source term
-    real(kind=8), parameter :: depth_tolerance = 1.0d-30
-
-    ! Friction source term
-    if (friction_forcing) then
-        do j=1,my
-            do i=1,mx
-                x = xlower + (i-0.5d0)*dx
-                ! Extract appropriate momentum
-                if (q(1,i,j) < depth_tolerance) then
-                    q(2:3,i,j) = 0.d0
-                else
-                    ! Apply friction source term only if in shallower water
-                    ! MODIFIED for benchmark 1 problem, drive to hu at inflow
-                    if (q(1,i,j) <= friction_depth) then
-                        do nman = num_manning, 1, -1
-                            if (aux(1,i,j) .lt. manning_break(nman)) then
-                                coeff = manning_coefficient(nman)
-                            endif
-                        enddo
-                        ! Calculate source term
-                        if (x < xforce) then
-                            gamma = 0.0055d0  ! based in inflow values, n=0.01
-                            dgamma = 1.d0 + dt * gamma
-                            !q2inflow = 0.054d0 * 0.115d0
-                            q2inflow = q(1,i,j) * 0.115d0  ! use h here
[…]

### 2015-03-02 dambreak_new/valout_vorticity.f

@@ -1,339 +1,0 @@
-c
-c -----------------------------------------------------
-c
-      subroutine valout (lst, lend, time, nvar, naux)
-c     -------------------------------------------------
-c
-c     Version that also outputs vorticity, so fort.q files have 5 columns:
-c         h, hu, hv, eta, vorticity
-c
-c
-      use amr_module
-      use geoclaw_module, only: dry_tolerance  
-      implicit double precision (a-h,o-z)
-      character*10  fname1, fname2, fname3, fname4, fname5
-
-c     # GeoClaw specific output....  add eta to q array before printing
-c
-c     # Write the results to the file fort.q<iframe>
-c     # Use format required by matlab script  plotclaw2.m or Python tools
-c
-c     # set outaux = .true. to also output the aux arrays to fort.a<iframe>
-
-      logical outaux
-      integer output_aux_num 
-      real(kind=8) :: h, hu, hv, eta
-      real(kind=8), allocatable :: qetav(:)
-      real(kind=8), allocatable :: u(:,:), v(:,:), vorticity(:,:)
-
-      integer :: clock_start, clock_finish, clock_rate
-
-      iadd(ivar,i,j)  = loc + ivar - 1 + nvar*((j-1)*mitot+i-1)
-      iaddaux(iaux,i,j) = locaux + iaux-1 + naux*(i-1) +
-     .                                      naux*mitot*(j-1)
-      iaddqetav(ivar,i,j)  = 1 + ivar - 1 + 5*((j-1)*mitot+i-1)
-c
-      call system_clock(clock_start,clock_rate)
-
-
-      if (nvar /= 3) then
-          write(6,*) '*** Error: valout assumes nvar==3 for geoclaw'
-          stop
-          endif
-
-c     # how many aux components requested?
-      output_aux_num = 0
-      do i=1,naux
-         output_aux_num = output_aux_num + output_aux_components(i)
-         enddo
-        
-c     # Currently outputs all aux components if any are requested!
-      outaux = ((output_aux_num > 0) .and. 
-     .         ((.not. output_aux_onlyonce) .or. (time==t0)))
-
-c     open(unit=77,file='fort.b',status='unknown',access='stream')
-
-
-c     ### Python graphics output
-c
-
-c        ###  make the file names and open output files
-         fname1 = 'fort.qxxxx'
-         fname2 = 'fort.txxxx'
-         fname3 = 'fort.axxxx'
[…]

### 2015-03-02 dambreak/maketopo.py
Added cylinder and square column in maketopo

@@ -8,49 +8,45 @@ def maketopo():
     """
     Output topography files
     """
-    nxpoints = 201
-    nypoints = 21
+    nxpoints = 260
+    nypoints = 12
     xlower = 0.
-    xupper = 20.
+    xupper = 16.6
     ylower = 0.
-    yupper = 1.52
+    yupper = 0.6
     outfile= "domain.tt1"
     topotools.topo1writer(outfile,topo,xlower,xupper,ylower,yupper,nxpoints,nypoints)
 
-    nxpoints = 201
-    nypoints = 201
-    xlower = 4.5
-    xupper = 5.5
-    ylower = 0.25
-    yupper = 1.25
-    outfile= "hump.tt1"
+    nxpoints = 260
+    nypoints = 12
+    xlower = 11.04
+    xupper = 11.16
+    ylower = 0.24
+    yupper = 0.36
+    outfile= "column.tt1"
     topotools.topo1writer(outfile,topo,xlower,xupper,ylower,yupper,nxpoints,nypoints)
 
 
 def topo(x,y):
     """
-    flat with hump
+    flat with cylinder
     """
 
     z = zeros(x.shape)
+    xo=11.1;
+    yo=0.3;
+    base_r=0.12/2;
 
-    # island addition
-    # from paper
-    xo=5.0;
-    yo=1.52/2;
-    slope_d=8;
-    base_r=0.75/2;
-    top_r=0.05/2;
-    hi=0.049;
-
-    dist = sqrt( (x-xo)**2+(y-yo)**2 );
-    zz = hi*(1-(dist-top_r)/(base_r-top_r))
-    zz = where(zz<hi, zz, hi)
-    z = where(dist<base_r, zz, 0.)
-
-    #z = z - 0.054
-    z = zeros(x.shape)
+    ## cylinder column
+#    hi=1.0;
+#    dist = sqrt( (x-xo)**2+(y-yo)**2 );
+#    zz = hi
+#    z = where(dist<base_r, zz, 0.)
     
+    ## square column
+    hi=1.0
+    z = where( ( (abs(x-xo)<base_r) & (abs(y-yo)<base_r) ),hi,0)
+
     return z
 
 def plot_topo():

### 2015-03-02 dambreak/setplot.py
Changed xlimits in plot

@@ -59,7 +59,7 @@ def setplot(plotdata):
     #plotaxes.axescmd = 'axes([.1,.7,.8,.2])'
     plotaxes.title = 'Surface'
     plotaxes.scaled = True
-    plotaxes.xlimits = [0,9.5]
+    plotaxes.xlimits = [0,16.6]
 
     cmap = colormaps.make_colormap({0:[0.5,1,0.5],0.01:[0,1,1], \
                                     0.2:[0,0,1], 0.5:[1,1,0],  0.8:[1,0,0], \
@@ -99,7 +99,7 @@ def setplot(plotdata):
     # Set up for axes in this figure:
     plotaxes = plotfigure.new_plotaxes()
     plotaxes.axescmd = 'subplot(312)'
-    plotaxes.xlimits = [0,12]
+    plotaxes.xlimits = [0,16.6]
     plotaxes.ylimits = [-0.05,0.5]
     plotaxes.title = 'Cross section of surface at y=0.3'
 
@@ -151,7 +151,7 @@ def setplot(plotdata):
 
     plotaxes = plotfigure.new_plotaxes()
     plotaxes.axescmd = 'subplot(313)'
-    plotaxes.xlimits = [0,12]
+    plotaxes.xlimits = [0,16.6]
     plotaxes.ylimits = [-0.2,2.0]
     plotaxes.title = 'u-velocity'
     plotitem = plotaxes.new_plotitem(plot_type='1d_from_2d_data')

### 2015-03-02 dambreak/setrun.py
Changed name of topo file

@@ -382,7 +382,7 @@ def setgeo(rundata):
     # for topography, append lines of the form
     #    [topotype, minlevel, maxlevel, t1, t2, fname]
     topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'domain.tt1'])
-    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'hump.tt1'])
+    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'column.tt1'])
 
     # == setdtopo.data values ==
     dtopo_data = rundata.dtopo_data

### 2015-03-02 dambreak/maketopo.py
Changed comment

@@ -29,7 +29,7 @@ def maketopo():
 
 def topo(x,y):
     """
-    flat with cylinder
+    flat with column
     """
 
     z = zeros(x.shape)

### 2015-03-02 dambreak/setplot.py
modified axis limit

@@ -59,7 +59,8 @@ def setplot(plotdata):
     #plotaxes.axescmd = 'axes([.1,.7,.8,.2])'
     plotaxes.title = 'Surface'
     plotaxes.scaled = True
-    plotaxes.xlimits = [0,9.5]
+    #plotaxes.xlimits = [0,9.5]
+    plotaxes.xlimits = 'auto' 
 
     cmap = colormaps.make_colormap({0:[0.5,1,0.5],0.01:[0,1,1], \
                                     0.2:[0,0,1], 0.5:[1,1,0],  0.8:[1,0,0], \
@@ -99,7 +100,8 @@ def setplot(plotdata):
     # Set up for axes in this figure:
     plotaxes = plotfigure.new_plotaxes()
     plotaxes.axescmd = 'subplot(312)'
-    plotaxes.xlimits = [0,12]
+    #plotaxes.xlimits = [0,12]
+    plotaxes.xlimits = 'auto'
     plotaxes.ylimits = [-0.05,0.5]
     plotaxes.title = 'Cross section of surface at y=0.3'
 
@@ -151,7 +153,8 @@ def setplot(plotdata):
 
     plotaxes = plotfigure.new_plotaxes()
     plotaxes.axescmd = 'subplot(313)'
-    plotaxes.xlimits = [0,12]
+    #plotaxes.xlimits = [0,12]
+    plotaxes.xlimits = 'auto'
     plotaxes.ylimits = [-0.2,2.0]
     plotaxes.title = 'u-velocity'
     plotitem = plotaxes.new_plotitem(plot_type='1d_from_2d_data')

### 2015-03-02 dambreak/maketopo.py
set topo to z=0 in maketopo

@@ -44,8 +44,8 @@ def topo(x,y):
 #    z = where(dist<base_r, zz, 0.)
     
     ## square column
-    hi=1.0
-    z = where( ( (abs(x-xo)<base_r) & (abs(y-yo)<base_r) ),hi,0)
+#    hi=1.0
+#    z = where( ( (abs(x-xo)<base_r) & (abs(y-yo)<base_r) ),hi,0)
 
     return z
 

### 2015-03-02 dambreak/setrun.py
let user choose topo file from setrun.py

@@ -379,10 +379,29 @@ def setgeo(rundata):
 
     # == settopo.data values ==
     topo_data = rundata.topo_data
+
+    #Set topography
+
+    # Choice of topography:
+    #   0 => flat bottom
+    #   1 => cylinder cloumn
+    #   2 => square column
+    topo_choice = 2
+
+    if (topo_choice == 0):
+        topo_domain_fname = 'domain_Zero.tt1'
+        topo_column_fname = 'columnZero.tt1'
+    elif(topo_choice == 1):
+        topo_domain_fname = 'domain_CylinderColumn.tt1'
+        topo_column_fname = 'columnCylinder.tt1'
+    elif(topo_choice == 2):
+        topo_domain_fname = 'domain_SquareColumn.tt1'
+        topo_column_fname = 'columnSquare.tt1'
+
     # for topography, append lines of the form
     #    [topotype, minlevel, maxlevel, t1, t2, fname]
-    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'domain.tt1'])
-    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'column.tt1'])
+    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_domain_fname])
+    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_column_fname])
 
     # == setdtopo.data values ==
     dtopo_data = rundata.dtopo_data

### 2015-03-02 dambreak/maketopo.py
modified maketopo

@@ -14,7 +14,7 @@ def maketopo():
     xupper = 16.6
     ylower = 0.
     yupper = 0.6
-    outfile= "domain.tt1"
+    outfile= "domain_CylinderColumn.tt1"
     topotools.topo1writer(outfile,topo,xlower,xupper,ylower,yupper,nxpoints,nypoints)
 
     nxpoints = 260
@@ -23,7 +23,7 @@ def maketopo():
     xupper = 11.16
     ylower = 0.24
     yupper = 0.36
-    outfile= "column.tt1"
+    outfile= "columnCylinder.tt1"
     topotools.topo1writer(outfile,topo,xlower,xupper,ylower,yupper,nxpoints,nypoints)
 
 
@@ -38,10 +38,10 @@ def topo(x,y):
     base_r=0.12/2;
 
     ## cylinder column
-#    hi=1.0;
-#    dist = sqrt( (x-xo)**2+(y-yo)**2 );
-#    zz = hi
-#    z = where(dist<base_r, zz, 0.)
+    hi=1.0;
+    dist = sqrt( (x-xo)**2+(y-yo)**2 );
+    zz = hi
+    z = where(dist<base_r, zz, 0.)
     
     ## square column
 #    hi=1.0

### 2015-03-02 dambreak/qinit.f90
changed final time

@@ -44,7 +44,7 @@ subroutine qinit(meqn,mbc,mx,my,xlower,ylower,dx,dy,q,maux,aux)
         enddo
     enddo
 
-!The grammar below seem to be incorrect for *.f90. It's only correct fo *.f
+!The grammar below seem to be incorrect for *.f90. It's only correct for *.f
 !    forall(i=1:mx, j=1:my)
 !        xcenter = xlower + (i - 0.5d0) * dx
 !        if (xcenter < sloc) then

### 2015-03-02 dambreak/setrun.py

@@ -119,8 +119,8 @@ def setrun(claw_pkg='geoclaw'):
 
     if clawdata.output_style==1:
         # Output nout frames at equally spaced times up to tfinal:
-        clawdata.num_output_times = 40
-        clawdata.tfinal = 12.0 
+        clawdata.num_output_times = 20
+        clawdata.tfinal = 17.0 
         clawdata.output_t0 = True  # output at initial (or restart) time?
 
     elif clawdata.output_style == 2:
@@ -332,8 +332,7 @@ def setrun(claw_pkg='geoclaw'):
     #gauges.append([2, x0+1.02, y0+0.27, 0., 1e10])
 
     gauges = rundata.gaugedata.gauges
-    gauges.append([1, 2.1, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
-    gauges.append([2, 11.1, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([1, 11.1, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
 
     return rundata
     # end of function setrun

### 2015-03-03 dambreak/myplot.py
add myplot.py for plot comparision between exp and simulation

@@ -1,0 +1,56 @@
+#!/usr/bin/python
+
+from clawpack.visclaw.data import ClawPlotData
+import matplotlib.pyplot as plt
+import os
+import csv
+import numpy as np
+
+
+plotdata = ClawPlotData()
+plotdata.outdir = '_output'   # set to the proper output directory
+gaugeno = 1                   # gauge number to examine
+g = plotdata.getgauge(gaugeno)
+#g.t is the array of times,
+#g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
+
+
+if not('myplot' in os.listdir('./')):
+    os.mkdir('myplot')
+
+#read experiment result and OpenFOAM result
+with open('./dataForComp/waveHeightNoColumnExp.csv', 'rb') as csvfile:
+    reader = csv.reader(csvfile)
+    waveHeightNoColumnExp = list(reader)
+    waveHeightNoColumnExp = np.array(waveHeightNoColumnExp)
+    waveHeightNoColumnExp = waveHeightNoColumnExp.astype(np.float)#convert string to floats
+    waveHeightNoColumnExp[:,1] = waveHeightNoColumnExp[:,1]+0.005751788#The exp data should be shifted before being used
+with open('./dataForComp/waveHeightNoColumnOF.csv', 'rb') as csvfile:
+    reader = csv.reader(csvfile)
+    waveHeightNoColumnOF = list(reader)
+    waveHeightNoColumnOF = np.array(waveHeightNoColumnOF)
+    waveHeightNoColumnOF = waveHeightNoColumnOF.astype(np.float)
+
+
+
+    
+
+
+
+#plot wave height history
+plt.figure()
+plt.plot(g.t,g.q[0,:],'b-.',waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',waveHeightNoColumnExp[:,0],waveHeightNoColumnExp[:,1],'g-')
+plt.xlabel('time (s)')
+plt.ylabel('wave height')
+plt.title('wave height history at x=11.1')
+plt.savefig('./myplot/waveheight.png', bbox_inches='tight')
+plt.close()
+
+#plot velocity u history
+plt.figure()
+plt.plot(g.t,g.q[1,:],'b-')
+plt.xlabel('time (s)')
+plt.ylabel('velocity u')
+plt.title('velocity u history at x=11.1')
+plt.savefig('./myplot/velocityU.png', bbox_inches='tight')
+plt.close()

### 2015-03-04 dambreak/myplot.py
modified myplot.py

@@ -8,10 +8,10 @@ import numpy as np
 
 
 plotdata = ClawPlotData()
-plotdata.outdir = '_output'   # set to the proper output directory
+plotdata.outdir = '_output_no_column'   # set to the proper output directory
 gaugeno = 1                   # gauge number to examine
 g = plotdata.getgauge(gaugeno)
-#g.t is the array of times,
+#g.t is the array of times
 #g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
 
 
@@ -31,6 +31,14 @@ with open('./dataForComp/waveHeightNoColumnOF.csv', 'rb') as csvfile:
     waveHeightNoColumnOF = np.array(waveHeightNoColumnOF)
     waveHeightNoColumnOF = waveHeightNoColumnOF.astype(np.float)
 
+with open('./dataForComp/velocity_nocolumn_exp.csv', 'rb') as csvfile:
+    reader = csv.reader(csvfile)
+    velocity_nocolumn_exp = list(reader)
+    velocity_nocolumn_exp = np.array(velocity_nocolumn_exp)
+    velocity_nocolumn_exp = velocity_nocolumn_exp.astype(np.float)
+
+#compute depth-averaged velocity u from experiment data
+
 
 
     
@@ -38,8 +46,12 @@ with open('./dataForComp/waveHeightNoColumnOF.csv', 'rb') as csvfile:
 
 
 #plot wave height history
+#plt.subplot(211)
 plt.figure()
-plt.plot(g.t,g.q[0,:],'b-.',waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',waveHeightNoColumnExp[:,0],waveHeightNoColumnExp[:,1],'g-')
+plt.plot(g.t,g.q[0,:],'b-.',label='geoclaw')
+plt.plot(waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',label='OF')
+plt.plot(waveHeightNoColumnExp[:,0],waveHeightNoColumnExp[:,1],'g--',label='Exp')
+legend = plt.legend(loc='upper center', shadow=True, fontsize='x-large')
 plt.xlabel('time (s)')
 plt.ylabel('wave height')
 plt.title('wave height history at x=11.1')

### 2015-03-04 dambreak/setplot.py
add gauges in setrun.py

@@ -61,6 +61,7 @@ def setplot(plotdata):
     plotaxes.scaled = True
     #plotaxes.xlimits = [0,9.5]
     plotaxes.xlimits = 'auto' 
+    #plotaxes.xlimits = [10.8,11.4]
 
 
     cmap = colormaps.make_colormap({0:[0.5,1,0.5],0.01:[0,1,1], \
@@ -103,6 +104,7 @@ def setplot(plotdata):
     plotaxes.axescmd = 'subplot(312)'
     #plotaxes.xlimits = [0,12]
     plotaxes.xlimits = 'auto'
+    #plotaxes.xlimits = [10.8,11.4]
     plotaxes.ylimits = [-0.05,0.5]
     plotaxes.title = 'Cross section of surface at y=0.3'
 
@@ -156,6 +158,7 @@ def setplot(plotdata):
     plotaxes.axescmd = 'subplot(313)'
     #plotaxes.xlimits = [0,12]
     plotaxes.xlimits = 'auto'
+    #plotaxes.xlimits = [10.8,11.4]
     plotaxes.ylimits = [-0.2,2.0]
     plotaxes.title = 'u-velocity'
     plotitem = plotaxes.new_plotitem(plot_type='1d_from_2d_data')

### 2015-03-04 dambreak/setrun.py

@@ -264,7 +264,7 @@ def setrun(claw_pkg='geoclaw'):
     amrdata = rundata.amrdata
 
     # max number of refinement levels:
-    amrdata.amr_levels_max = 1
+    amrdata.amr_levels_max = 2
 
     # List of refinement ratios at each level (length at least mxnest-1)
     amrdata.refinement_ratios_x = [8]
@@ -332,7 +332,12 @@ def setrun(claw_pkg='geoclaw'):
     #gauges.append([2, x0+1.02, y0+0.27, 0., 1e10])
 
     gauges = rundata.gaugedata.gauges
-    gauges.append([1, 11.1, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([1, 11.16, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([2, 11.04, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([3, 11.03, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([4, 11.02, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([5, 11.01, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
+    gauges.append([6, 11.00, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
 
     return rundata
     # end of function setrun

### 2015-03-04 doc/paper/paper.tex

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amssymb}

\usepackage{hyperref}  % for urls and hyperlinks

\setlength{\textwidth}{6.2in}
\setlength{\oddsidemargin}{0.3in}
\setlength{\evensidemargin}{0in}
\setlength{\textheight}{8.7in}
\setlength{\voffset}{-.7in}
\setlength{\headsep}{26pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

% input some useful macros from RJLmacros.tex:
\input{../macros/RJLmacros}

\title{Comparison of the solution of the shallow equation with experimental}

\author{Xinsheng Qin xsqin@uw.edu\\
        Kaspar Mueller kasparm@uw.edu}

\begin{document}

\maketitle

\begin{abstract}
In this paper we compare the solution of the shallow water equations with experimental results. The experiment simulates the interaction between an incident bore and a free-standing coastal structure. The shallow water equations are solved using the CLAWPACK software.
\end{abstract}

\section{Introduction}\label{Sec:intro}
The aim of this work is to compare computational results using CLAWPACK with experimental results. This is done for the interaction of a bore and a free-standing coastal structure. The experiment results are given by Halld\'or \'Arnason \cite{HA}. We model the water using the shallow water equations which will be introduced in Section \ref{Sec:ShallowWater}, followed by a description how CLAWPACK solves the system and handles the mesh in Section \ref{Sec:Algo}. In Section \ref{Sec:Setup} we describe the setup of the test case. We compare wave height and velocity for the case of a pure dambreak without structure in the domain in Section \ref{Sec:Zero}. In Section \ref{Sec:Square} we add a square column and compare wave height upstream as well as downstream of the column. In Section \ref{Sec:Cylinder} we compare the wave height at the same location but for a cylinder column. We close this paper with a conclusion in Section \ref{Sec:Conclusion}.
\section{The shallow water equations}\label{Sec:ShallowWater}
\eqm
h_t + (uh)_x + (vh)_y = 0\\
(hu)_t + (huv)_y +(hu^2 + \frac{1}{2}gh^2)_x = -ghB_x - Du\\
(hv)_t + (huv)_x +(hv^2 + \frac{1}{2}gh^2)_y = -ghB_y - Dv
\enm
where $u,v$ are the horizontal depth averaged velocity, $B$ the topography and $D$ the drag coefficient.
\section{The Algorithm in CLAWPACK/GeoClaw}\label{Sec:Algo}
We solve the shallow water equations with CLAWPACK. Specifically we used the GeoClaw solver which used the f-wave approach described in \cite{DLG}. There the difference in flux normal to the cell interface is separated into propagating waves rather than the jump in $Q$.
\section{The setup of the test case}\label{Sec:Setup}
The benchmark experiment was conducted in a 16.6 m long, 0.6m wide and 0.45 m deep wave tank. When a gate is lifted, it generates a single bore to impinge the column located downstream.
We are going to simulate 2 cases in the experiment. The 1st case purely generated the single bore, the 2nd case generated the bore to impinge the column with one side facing the flow.
\par
In the 1st case, wave height and streamwise velocity, U, at different depth, at where the center of column should be located are measured. In the 2nd case, the wave height in front and behind the column are measured.
        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.5\textwidth]{../proposal/figures/Diagram_case}
            \caption{Diagram of experiment apparatus}
        \end{figure}
\section{Dam Break}\label{Sec:Zero}
\section{Dam Break with square column}\label{Sec:Square}
\section{Dam Break with cylinder column}\label{Sec:Cylinder}
\section{Conclusion}\label{Sec:Conclusion}
{\footnotesize
\begin{thebibliography}{100}
\bibitem{HA} Halld\'or \'Arnason  (2005). Thesis - Interactions between an incident bore and a free-standing coastal structure. \url{http://faculty.washington.edu/cpetroff/Halldor%20dissertation.pdf}
\bibitem{DLG} David L. George (2008) Journal of Computational Physics - Augmented Riemann solvers for the shallow water equations over variable topography with steady states and inundation.
\bibitem{MJB} Marsha J. Berger and David L. George and Randall J. LeVeque and Kyle T. Mandli (2011) Advances in Water Resources - The GeoClaw software for depth-averaged flows with adaptive refinement.
\end{thebibliography}
}
\end{document}

### 2015-03-04 doc/paper/paper.tex
Some more text added to draft

@@
 \section{The shallow water equations}\label{Sec:ShallowWater}
+We write the shallow water equations as
 \eqm
@@
 \enm
-where $u,v$ are the horizontal depth averaged velocity, $B$ the topography and $D$ the drag coefficient.
+where $u,v$ are the horizontal depth averaged velocities, $B$ the topography and $D$ the drag coefficient.
+$g$ denotes the gravitational acceleration.
+In the present case the topography will be flat except the column located downstream.
+The drag coefficient $D$ is given by
+\eq
+D = \frac{gM^2\sqrt{(u^2+v^2)}}{h^{5/3}}
+\en
+where $M$ is the Manning coefficient which is $M = 0.015$.
+The Manning coefficient is an empirical coefficient that includes the roughness of the surface.
+The initial condition is given by a jump in the height $h$.
+The water is initially at rest.
 \section{The Algorithm in CLAWPACK/GeoClaw}\label{Sec:Algo}
@@
 Specifically we used the GeoClaw solver which used the f-wave approach described in \cite{DLG}.
-There the difference in flux normal to the cell interface is separated into propagating waves rather than the jump in $Q$.
+There the difference in flux normal to the cell interface is separated into propagating waves.
+The topography and drag term are incorporated in the Riemann problem.
+Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.
 \section{The setup of the test case}\label{Sec:Setup}
@@
 \section{Dam Break}\label{Sec:Zero}
+The initial condition for this and the following cases is given by
+\eqm
+h_0(x,y)=& 
+\begin{cases}
+0.25 & x<5.9 \\
+0.02 & \text{else}
+\end{cases}\\
+u_0(x,y) =& 0
+\enm
 \section{Dam Break with square column}\label{Sec:Square}

### 2015-03-05 dambreak/myplot.py
modified myplot.py

@@ -6,10 +6,12 @@ import os
 import csv
 import numpy as np
 
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+#for case without a column
 
 plotdata = ClawPlotData()
-plotdata.outdir = '_output_no_column'   # set to the proper output directory
-gaugeno = 1                   # gauge number to examine
+plotdata.outdir = '_output_nocolumn'   # set to the proper output directory
+gaugeno = 7                   # gauge number to examine
 g = plotdata.getgauge(gaugeno)
 #g.t is the array of times
 #g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
@@ -37,32 +39,99 @@ with open('./dataForComp/velocity_nocolumn_exp.csv', 'rb') as csvfile:
     velocity_nocolumn_exp = np.array(velocity_nocolumn_exp)
     velocity_nocolumn_exp = velocity_nocolumn_exp.astype(np.float)
 
-#compute depth-averaged velocity u from experiment data
-
-
-
-    
-
-
 
 #plot wave height history
-#plt.subplot(211)
 plt.figure()
 plt.plot(g.t,g.q[0,:],'b-.',label='geoclaw')
-plt.plot(waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',label='OF')
+plt.plot(waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',label='OpenFOAM')
 plt.plot(waveHeightNoColumnExp[:,0],waveHeightNoColumnExp[:,1],'g--',label='Exp')
 legend = plt.legend(loc='upper center', shadow=True, fontsize='x-large')
 plt.xlabel('time (s)')
-plt.ylabel('wave height')
+plt.ylabel('wave height (m)')
 plt.title('wave height history at x=11.1')
-plt.savefig('./myplot/waveheight.png', bbox_inches='tight')
+plt.savefig('./myplot/waveheight_nocolumn.png', bbox_inches='tight')
 plt.close()
 
 #plot velocity u history
 plt.figure()
-plt.plot(g.t,g.q[1,:],'b-')
+#q[1] is hu instead of u
+u=np.zeros(g.t.size)
+for i in range(g.t.size):
+    u[i] = g.q[1,i]/g.q[0,i]
[…]

### 2015-03-05 doc/paper/paper.tex
modified paper.tex

--- 2015-03-04 doc/paper/paper.tex
@@
         \end{figure}
+
+In all cases below, boundary conditions are all set to reflection wall. 
 \section{Dam Break}\label{Sec:Zero}
+The initial condition for this and the following cases is given by
+\eqm
+h_0(x,y)=& 
+\begin{cases}
+0.25 & x<5.9 \\
+0.02 & \text{else}
+\end{cases}\\
+u_0(x,y) =& 0
+\enm
+
+Figure \ref{fig:waveheight_nocolumn} shows time history of wave height measured at where center of column should be. 
+Result from geoclaw agree quite well in general. At around t=3.8, geoclaw accurately predicted the arrival of the shock. 
+As for the peak value of this shock, geoclaw overestimated it a little, while OpenFOAM, another open source computational fluid dynamics software gave a better value by solving N-S equations with turbulent model.
+After a long time, the wave hit ending wall and got reflected, which can be told by the second jump in wave height in the figure. 
+Geoclaw also predicted the arrival time of this peak quite well and got only a little overestimation of peak value.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/waveheight_nocolumn}
+    \caption{time history of wave height for case without a column}
+    \label{fig:waveheight_nocolumn}
+\end{figure}
+
+Figure \ref{fig:velocityU_nocolumn} gives time history of streamwise velocity u from geoclaw and experiment. 
+Note that this velocity u in both geoclaw and experiment are depth-averaged. 
+Geoclaw accurately predicted the arrival time of velocity peak as well as value of the peak. 
+After the peak arrived, geoclaw overestimated value of velocity.
+The reason why geoclaw gave a higher velocity may be attributed to its neglection of fraction force from tank bottom. In experiment, velocity get smaller as it comes closer to the bottom. 
+In geoclawa, however, there is only one u in depth direction and it does not take fraction into account. 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/velocityU_nocolumn}
+    \caption{time history of streamwise velocity U for case without a column}
+    \label{fig:velocityU_nocolumn}
+\end{figure}
+
 \section{Dam Break with square column}\label{Sec:Square}
+Figure \ref{fig:waveheight_square_x=11.0} gives comparison of time history of wave height measured at 0.04m ahead of the leading edge of square column.(0.1m ahead of center of the column)
+Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to previous figures.
+Arrival time of peak predicted by geoclaw is about 0.2 second earlier than time measured from experiment. 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/waveheight_square_x=11}
+    \caption{time history of wave height for case with square column}
+    \label{fig:waveheight_square_x=11.0}
+\end{figure}
+
 \section{Dam Break with cylinder column}\label{Sec:Cylinder}
+Figure \ref{fig:waveheight_cylinder_x=11.0} is time history of wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of center of the column)
+Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
+Thus the result in this case does not agree well with experiment result.
+Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/waveheight_cylinder_x=11}
+    \caption{time history of wave height for case with cylinder column}
+    \label{fig:waveheight_cylinder_x=11.0}
+\end{figure}
+
 \section{Conclusion}\label{Sec:Conclusion}

### 2015-03-05 doc/paper/paper.tex
Some changes to the paper draft

@@
 The Manning coefficient is an empirical coefficient that includes the roughness of the surface.
-The initial condition is given by a jump in the height $h$.
+The initial condition is given by a jump in the water depth $h$.
 The water is initially at rest.
@@
 
-Figure \ref{fig:waveheight_nocolumn} shows time history of wave height measured at where center of column should be. 
+Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height measured at the center location where the column will be mounted. 
 Result from geoclaw agree quite well in general. At around t=3.8, geoclaw accurately predicted the arrival of the shock. 
-As for the peak value of this shock, geoclaw overestimated it a little, while OpenFOAM, another open source computational fluid dynamics software gave a better value by solving N-S equations with turbulent model.
-After a long time, the wave hit ending wall and got reflected, which can be told by the second jump in wave height in the figure. 
-Geoclaw also predicted the arrival time of this peak quite well and got only a little overestimation of peak value.
+As for the peak value of this shock, geoclaw overestimated it a little, while OpenFOAM, another open source computational fluid dynamics software gave a better value by solving the Navier-Stokes equations with turbulent model.
+After a longer time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height in the figure. 
+Geoclaw also predicted the arrival time of this peak quite well and got only a little overestimation of the peak value.
 \begin{figure}[h!]
@@
 
-Figure \ref{fig:velocityU_nocolumn} gives time history of streamwise velocity u from geoclaw and experiment. 
-Note that this velocity u in both geoclaw and experiment are depth-averaged. 
-Geoclaw accurately predicted the arrival time of velocity peak as well as value of the peak. 
-After the peak arrived, geoclaw overestimated value of velocity.
-The reason why geoclaw gave a higher velocity may be attributed to its neglection of fraction force from tank bottom. In experiment, velocity get smaller as it comes closer to the bottom. 
+Figure \ref{fig:velocityU_nocolumn} gives the time history of the streamwise velocity $u$ from geoclaw and experiment. 
+Note that this velocity $u$ in both geoclaw and experiment are depth-averaged. 
+Geoclaw accurately predicted the time of the velocity peak as well as the value of the peak. 
+After the peak arrived, geoclaw overestimated the value of velocity.
+(The reason why geoclaw gives a higher velocity may be attributed to its neglection of fraction force from the tank bottom).
+In the experiment, the velocity gets smaller as it comes closer to the bottom.
 In geoclawa, however, there is only one u in depth direction and it does not take fraction into account. 
@@
 \section{Dam Break with square column}\label{Sec:Square}
-Figure \ref{fig:waveheight_square_x=11.0} gives comparison of time history of wave height measured at 0.04m ahead of the leading edge of square column.(0.1m ahead of center of the column)
-Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to previous figures.
-Arrival time of peak predicted by geoclaw is about 0.2 second earlier than time measured from experiment. 
+Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at 0.04m ahead of the leading edge of the square column.(0.1m ahead of center of the column)
+Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to the previous figures.
+The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment. 
 \begin{figure}[h!]
@@
 \section{Dam Break with cylinder column}\label{Sec:Cylinder}
-Figure \ref{fig:waveheight_cylinder_x=11.0} is time history of wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of center of the column)
+Figure \ref{fig:waveheight_cylinder_x=11.0} is the time history of the wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of the center of the column)
 Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
-Thus the result in this case does not agree well with experiment result.
+Thus the result in this case does not agree well with the experiment result.
 Future test can be conducted by setting a higher level for mesh refinement to get a better circle.

### 2015-03-05 dambreak/myplot.py
modified myplot.py

@@ -8,130 +8,177 @@ import numpy as np
 
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 #for case without a column
-
-plotdata = ClawPlotData()
-plotdata.outdir = '_output_nocolumn'   # set to the proper output directory
-gaugeno = 7                   # gauge number to examine
-g = plotdata.getgauge(gaugeno)
-#g.t is the array of times
-#g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
-
-
-if not('myplot' in os.listdir('./')):
-    os.mkdir('myplot')
-
-#read experiment result and OpenFOAM result
-with open('./dataForComp/waveHeightNoColumnExp.csv', 'rb') as csvfile:
-    reader = csv.reader(csvfile)
-    waveHeightNoColumnExp = list(reader)
-    waveHeightNoColumnExp = np.array(waveHeightNoColumnExp)
-    waveHeightNoColumnExp = waveHeightNoColumnExp.astype(np.float)#convert string to floats
-    waveHeightNoColumnExp[:,1] = waveHeightNoColumnExp[:,1]+0.005751788#The exp data should be shifted before being used
-with open('./dataForComp/waveHeightNoColumnOF.csv', 'rb') as csvfile:
-    reader = csv.reader(csvfile)
-    waveHeightNoColumnOF = list(reader)
-    waveHeightNoColumnOF = np.array(waveHeightNoColumnOF)
-    waveHeightNoColumnOF = waveHeightNoColumnOF.astype(np.float)
-
-with open('./dataForComp/velocity_nocolumn_exp.csv', 'rb') as csvfile:
-    reader = csv.reader(csvfile)
-    velocity_nocolumn_exp = list(reader)
-    velocity_nocolumn_exp = np.array(velocity_nocolumn_exp)
-    velocity_nocolumn_exp = velocity_nocolumn_exp.astype(np.float)
+def plot_no_column():
+    plotdata = ClawPlotData()
+    plotdata.outdir = '_output_nocolumn'   # set to the proper output directory
+    gaugeno = 7                   # gauge number to examine
+    g = plotdata.getgauge(gaugeno)
+    #g.t is the array of times
[…]

### 2015-03-05 dambreak/setrun.py
modified setrun.py

@@ -71,8 +71,8 @@ def setrun(claw_pkg='geoclaw'):
 
 
     # Number of grid cells: Coarsest grid
-    clawdata.num_cells[0] = 332
-    clawdata.num_cells[1] = 12
+    clawdata.num_cells[0] = 830 
+    clawdata.num_cells[1] = 30 
 
 
     # ---------------
@@ -267,9 +267,9 @@ def setrun(claw_pkg='geoclaw'):
     amrdata.amr_levels_max = 2
 
     # List of refinement ratios at each level (length at least mxnest-1)
-    amrdata.refinement_ratios_x = [8]
-    amrdata.refinement_ratios_y = [8]
-    amrdata.refinement_ratios_t = [8]
+    amrdata.refinement_ratios_x = [2]
+    amrdata.refinement_ratios_y = [2]
+    amrdata.refinement_ratios_t = [2]
 
 
     # Specify type of each aux variable in amrdata.auxtype.
@@ -317,8 +317,8 @@ def setrun(claw_pkg='geoclaw'):
     regions = rundata.regiondata.regions
     # to specify regions of refinement append lines of the form
     #  [minlevel,maxlevel,t1,t2,x1,x2,y1,y2]
-    regions.append([1, 1, 0., 1e9, 0., 100., 0., 2.])
-    regions.append([2, 2, 0., 1e9, 3., 10., 0., 2.])
+    regions.append([1, 1, 0., 1e9, 0., 16.6, 0., 0.6])
+    regions.append([2, 2, 0., 1e9, 10.8, 11.4, 0.2, 0.4])
 
     # == setgauges.data values ==
     # for gauges append lines of the form  [gaugeno, x, y, t1, t2]
@@ -404,7 +404,8 @@ def setgeo(rundata):
 
     # for topography, append lines of the form
     #    [topotype, minlevel, maxlevel, t1, t2, fname]
-    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_domain_fname])
+    #topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_domain_fname])
+    topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'domain_Zero.tt1'])
     topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_column_fname])
 
     # == setdtopo.data values ==

### 2015-03-06 dambreak/setrun.py
Cleaned up topo files

@@ -393,18 +393,14 @@ def setgeo(rundata):
     topo_choice = 2
 
     if (topo_choice == 0):
-        topo_domain_fname = 'domain_Zero.tt1'
         topo_column_fname = 'columnZero.tt1'
     elif(topo_choice == 1):
-        topo_domain_fname = 'domain_CylinderColumn.tt1'
         topo_column_fname = 'columnCylinder.tt1'
     elif(topo_choice == 2):
-        topo_domain_fname = 'domain_SquareColumn.tt1'
         topo_column_fname = 'columnSquare.tt1'
 
     # for topography, append lines of the form
     #    [topotype, minlevel, maxlevel, t1, t2, fname]
-    #topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_domain_fname])
     topo_data.topofiles.append([1, 1, 1, 0., 1.e10, 'domain_Zero.tt1'])
     topo_data.topofiles.append([1, 1, 1, 0., 1.e10, topo_column_fname])
 

### 2015-03-06 dambreak/myplot.py
myplot.py and setrun.py

@@ -118,7 +118,7 @@ def plot_cylinder():
     legend = plt.legend(loc='lower center', shadow=True, fontsize='x-large')
     plt.xlabel('time (s)')
     plt.ylabel('wave height (m)')
-    plt.title('wave height history at x=11.0')
+    plt.title('wave height history at x=11.16')
     plt.xlim(3.0,4.0)
     plt.savefig('./myplot/waveheight_cylinder_x=11.16.png', bbox_inches='tight')
     plt.close()
@@ -129,11 +129,19 @@ def plot_square():
     plotdata = ClawPlotData()
     plotdata.outdir = '_output_square'   # set to the proper output directory
     gaugeno = 6                   # gauge number to examine
-    g = plotdata.getgauge(gaugeno)
+    g1 = plotdata.getgauge(gaugeno)
     #g.t is the array of times
     #g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
 
+    plotdata = ClawPlotData()
+    plotdata.outdir = '_output_square_2'   # set to the proper output directory
+    gaugeno = 6                   # gauge number to examine
+    g2 = plotdata.getgauge(gaugeno)
 
+    plotdata = ClawPlotData()
+    plotdata.outdir = '_output_square_3'   # set to the proper output directory
+    gaugeno = 6                   # gauge number to examine
+    g3 = plotdata.getgauge(gaugeno)
     if not('myplot' in os.listdir('./')):
         os.mkdir('myplot')
 
@@ -147,7 +155,9 @@ def plot_square():
 
     #plot waveheight history
     plt.figure()
-    plt.plot(g.t,g.q[0,:],'b-',label='geoclaw')
+    plt.plot(g1.t,g1.q[0,:],'b-',label='geoclaw1')
+    plt.plot(g2.t,g2.q[0,:],'g-.',label='geoclaw2')
+    plt.plot(g3.t,g3.q[0,:],'y-o',label='geoclaw3')
     plt.plot(wave_height_square[:,0],wave_height_square[:,1],'r--',label='Exp')
     legend = plt.legend(loc='lower center', shadow=True, fontsize='x-large')
     plt.xlabel('time (s)')
@@ -156,10 +166,23 @@ def plot_square():
     plt.xlim(3.0,4.0)
     plt.savefig('./myplot/waveheight_square_x=11.0.png', bbox_inches='tight')
     plt.close()
+
[…]

### 2015-03-06 dambreak/setrun.py

@@ -163,7 +163,7 @@ def setrun(claw_pkg='geoclaw'):
 
     # Initial time step for variable dt.
     # If dt_variable==0 then dt=dt_initial for all steps:
-    clawdata.dt_initial = 0.016
+    clawdata.dt_initial = 0.0016
 
     # Max time step to be allowed if variable dt used:
     clawdata.dt_max = 1e+99
@@ -264,12 +264,15 @@ def setrun(claw_pkg='geoclaw'):
     amrdata = rundata.amrdata
 
     # max number of refinement levels:
-    amrdata.amr_levels_max = 2
+    amrdata.amr_levels_max = 4
 
     # List of refinement ratios at each level (length at least mxnest-1)
-    amrdata.refinement_ratios_x = [2]
-    amrdata.refinement_ratios_y = [2]
-    amrdata.refinement_ratios_t = [2]
+    #amrdata.refinement_ratios_x = [2,2]
+    #amrdata.refinement_ratios_y = [2,2]
+    #amrdata.refinement_ratios_t = [2,2]
+    amrdata.refinement_ratios_x = [2,2,2]
+    amrdata.refinement_ratios_y = [2,2,2]
+    amrdata.refinement_ratios_t = [2,2,2]
 
 
     # Specify type of each aux variable in amrdata.auxtype.
@@ -318,7 +321,7 @@ def setrun(claw_pkg='geoclaw'):
     # to specify regions of refinement append lines of the form
     #  [minlevel,maxlevel,t1,t2,x1,x2,y1,y2]
     regions.append([1, 1, 0., 1e9, 0., 16.6, 0., 0.6])
-    regions.append([2, 2, 0., 1e9, 10.8, 11.4, 0.2, 0.4])
+    regions.append([3, 4, 0., 1e9, 10.8, 11.4, 0.2, 0.4])
 
     # == setgauges.data values ==
     # for gauges append lines of the form  [gaugeno, x, y, t1, t2]
@@ -332,8 +335,8 @@ def setrun(claw_pkg='geoclaw'):
     #gauges.append([2, x0+1.02, y0+0.27, 0., 1e10])
 
     gauges = rundata.gaugedata.gauges
-    gauges.append([1, 11.16, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
-    gauges.append([2, 11.04, 0.3, 0., 1e10]) #This values of t1 and t2 specified means that this gauge data will be output for all times
[…]

### 2015-03-07 dambreak/myplot.py
modified myplot.py

@@ -11,7 +11,7 @@ import numpy as np
 def plot_no_column():
     plotdata = ClawPlotData()
     plotdata.outdir = '_output_nocolumn'   # set to the proper output directory
-    gaugeno = 7                   # gauge number to examine
+    gaugeno = 6                   # gauge number to examine
     g = plotdata.getgauge(gaugeno)
     #g.t is the array of times
     #g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
@@ -46,10 +46,11 @@ def plot_no_column():
     plt.plot(waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',label='OpenFOAM')
     plt.plot(waveHeightNoColumnExp[:,0],waveHeightNoColumnExp[:,1],'g--',label='Exp')
     legend = plt.legend(loc='upper center', shadow=True, fontsize='x-large')
+    plt.xlim(3.0,4.0)
     plt.xlabel('time (s)')
     plt.ylabel('wave height (m)')
     plt.title('wave height history at x=11.1')
-    plt.savefig('./myplot/waveheight_nocolumn.png', bbox_inches='tight')
+    plt.savefig('./myplot/waveheight_nocolumn_xlim_x=11.0.png', bbox_inches='tight')
     plt.close()
 
     #plot velocity u history
@@ -75,9 +76,12 @@ def plot_cylinder():
     plotdata = ClawPlotData()
     plotdata.outdir = '_output_cylinder_2'   # set to the proper output directory
     gaugeno = 6                   # gauge number to examine
-    g = plotdata.getgauge(gaugeno)
-    #g.t is the array of times
-    #g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
+    g1 = plotdata.getgauge(gaugeno)
+
+    plotdata = ClawPlotData()
+    plotdata.outdir = '_output_finerColumntt1'   # set to the proper output directory
+    gaugeno = 6                   # gauge number to examine
+    g2 = plotdata.getgauge(gaugeno)
     if not('myplot' in os.listdir('./')):
         os.mkdir('myplot')
     #read experiment result and OpenFOAM result
[…]

### 2015-03-09 doc/paper/paper.tex
add some contents in paper.tex

@@
 After the peak arrived, geoclaw overestimated the value of velocity.
-(The reason why geoclaw gives a higher velocity may be attributed to its neglection of fraction force from the tank bottom).
+The reason why geoclaw gives a higher velocity may be attributed to its depth-averaged property: water at all depth are propagating with the same speed.
 In the experiment, the velocity gets smaller as it comes closer to the bottom.
-In geoclawa, however, there is only one u in depth direction and it does not take fraction into account. 
 \begin{figure}[h!]
@@
 Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to the previous figures.
-The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment. 
+There are two jump in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
+If we recall that the wave height is around 0.1 in \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak.
+The second jump is induced by wave reflected by the column.
+%todo
+If we compare the time interval between first jump and second jump with that in \ref{}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
+%
+The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
+This can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
+In shallow water equations, fluid viscosity is not taken into account, while it's significant in N-S equation (or real physics).
+This causes the shock remain its shape in as it propagates in shallow water equation and smooth out in experiment.
+Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
+Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
+Thus before the propagating shock arrive, all water downstream the shock stays still.
+In other words, it even does not know there is a column before the shock hits the column.
+In contrast, N-S equation is a elliptic PDE.
+In real physics govened by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent, which raises the wave hegiht even before the wave hits the column.
+This explains why we cannot see two jumps in experiment results.
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/waveheight_square_x=11}
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x=11}
     \caption{time history of wave height for case with square column}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/waveheight_cylinder_x=11}
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x=11}
     \caption{time history of wave height for case with cylinder column}

### 2015-03-09 doc/paper/paper.tex
Fixed figure file naming and location
add a plots in paper.tex

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/waveheight_nocolumn}
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn}
     \caption{time history of wave height for case without a column}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{../../dambreak/myplot/velocityU_nocolumn}
+    \includegraphics[width=0.5\textwidth]{./plots/velocityU_nocolumn}
     \caption{time history of streamwise velocity U for case without a column}
@@
 Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to the previous figures.
-There are two jump in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
+There are two jumps in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
 If we recall that the wave height is around 0.1 in \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak.
 The second jump is induced by wave reflected by the column.
-%todo
-If we compare the time interval between first jump and second jump with that in \ref{}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
-%
+If we compare the time interval between first jump and second jump with that in \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
 The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x=11}
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x11}
     \caption{time history of wave height for case with square column}
     \label{fig:waveheight_square_x=11.0}
+\end{figure}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
+    \caption{time history of wave height for case with square column, at x = 11.02}
+    \label{fig:waveheight_square_x=11.02}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x=11}
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x11}
     \caption{time history of wave height for case with cylinder column}

### 2015-03-09 doc/paper/paper.tex
Changes in first few paragraphs

--- 2015-03-09 doc/paper/paper.tex
@@
 The experiment results are given by Halld\'or \'Arnason \cite{HA}.
-We model the water using the shallow water equations which will be introduced in Section \ref{Sec:ShallowWater}, followed by a description how CLAWPACK solves the system and handles the mesh in Section \ref{Sec:Algo}.
+We model the water using the shallow water equations which will be introduced in Section \ref{Sec:ShallowWater}.
+We solve these equations using CLAWPACK's GeoClaw solver which is specialized for geophysical flow problems.
+In Section \ref{Sec:Algo} we show how GeoClaw solves the shallow water equations and how it handles the topography.
 In Section \ref{Sec:Setup} we describe the setup of the test case.
-We compare wave height and velocity for the case of a pure dambreak without structure in the domain in Section \ref{Sec:Zero}.
-In Section \ref{Sec:Square} we add a square column and compare wave height upstream as well as downstream of the column.
-In Section \ref{Sec:Cylinder} we compare the wave height at the same location but for a cylinder column.
+We focus on three cases.
+First we compare wave height and velocity for the pure dambreak without structure in the domain in Section \ref{Sec:Zero}.
+Second we add a square column and compare wave height upstream as well as downstream of the column in \ref{Sec:Square}.
+Lastly in Section \ref{Sec:Cylinder} we compare the wave height at the same location but for a cylinder column.
 We close this paper with a conclusion in Section \ref{Sec:Conclusion}.
 \section{The shallow water equations}\label{Sec:ShallowWater}
-We write the shallow water equations as
+We write the 2D depth averaged shallow water equations as
 \eqm
@@
 We solve the shallow water equations with CLAWPACK.
-Specifically we used the GeoClaw solver which used the f-wave approach described in \cite{DLG}.
-There the difference in flux normal to the cell interface is separated into propagating waves.
+Specifically we used the GeoClaw solver which used a special version of the f-wave approach described in \cite{DLG}.
+This solver combines the qualities of the Roe solver, HLLE-type solver and the f-wave approach.
+From the Roe solver it has the quality of giving a exact solution for a single-shock Riemann problem.
+Like the HLLE solver this method is depth positive semidefinite.
+The method has also a natural entropy-fix and gives a better approximation for problems with large rarefactions.
+The solver also maintains large class of steady states.
 The topography and drag term are incorporated in the Riemann problem.
-Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.
+Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.\\
+In the GeoClaw solver adaptive mesh refinement (AMR) can be used.
+Multiple levels can be use where for each level the domain is specified.
+The topography for the problem is defined in a file.
+The topography goes into the shallow water equations through the right hand side $B$.
+The value from the file are converted into cell averages using a picewise bilinear function that interpolates the pointwise values.
+$B$ is then obtained by evaluating the exact integral of this function over the cell.
 \section{The setup of the test case}\label{Sec:Setup}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x=11}
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x11}
     \caption{time history of wave height for case with square column}

### 2015-03-09 doc/paper/paper.tex

@@
     \label{fig:waveheight_square_x=11.0}
+\end{figure}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
+    \caption{time history of wave height for case with square column, at x = 11.02}
+    \label{fig:waveheight_square_x=11.02}
 \end{figure}

### 2015-03-09 doc/paper/paper.tex
some modification on paper.tex

--- 2015-03-09 doc/paper/paper.tex
@@
     \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn}
-    \caption{time history of wave height for case without a column}
+    \caption{time history of wave height for case without a column. }
     \label{fig:waveheight_nocolumn}
@@
 If we compare the time interval between first jump and second jump with that in \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
-The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
-This can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
+\par
+The arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
+If we zoom in \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get \ref{fig:waveheight_nocolumn_zoomin}.
+In \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is about 0.11 second earlier that that from experiment.  
+The difference in shock shape can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
 In shallow water equations, fluid viscosity is not taken into account, while it's significant in N-S equation (or real physics).
-This causes the shock remain its shape in as it propagates in shallow water equation and smooth out in experiment.
+This causes the shock remain its shape in as it propagates in shallow water equation while smooth out in experiment.
 Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
+\par
 Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
 Thus before the propagating shock arrive, all water downstream the shock stays still. 
-In other words, it even does not know there is a column before the shock hits the column.
-In contrast, N-S equation is a elliptic PDE.
+In other words, it even does not know there is a column before the shock hits the column. In contrast, N-S equation is a elliptic PDE. 
 In real physics govened by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent, which raises the wave hegiht even before the wave hits the column.
@@
     \label{fig:waveheight_square_x=11.02}
+\end{figure}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
+    \caption{time history of wave height for case with square column, at x = 11.02}
+    \label{fig:waveheight_nocolumn_zoomin}
 \end{figure}

### 2015-03-09 doc/paper/paper.tex
Changes in part DamBreak

--- 2015-03-09 doc/paper/paper.tex
@@
 0.25 & x<5.9 \\
-0.02 & \text{else}
+0.02 & \text{otherwise}
 \end{cases}\\
@@
 \enm
-
+In this comparison we will include results from OpenFOAM, an open source CFD software modeling air and water using the Navier-Stokes equations with turbulence model.
 Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height measured at the center location where the column will be mounted. 
-Result from geoclaw agree quite well in general. At around t=3.8, geoclaw accurately predicted the arrival of the shock. 
-As for the peak value of this shock, geoclaw overestimated it a little, while OpenFOAM, another open source computational fluid dynamics software gave a better value by solving the Navier-Stokes equations with turbulent model.
-After a longer time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height in the figure. 
-Geoclaw also predicted the arrival time of this peak quite well and got only a little overestimation of the peak value.
+Result from GeoClaw agree quite well in general(Show zoom in plot). At around t=3.8, GeoClaw accurately predicted the arrival of the shock. 
+GeoClaw overestimates the peak water level a little, while OpenFOAM, gives a better value.
+After a later time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height. 
+GeoClaw also predicts the arrival time of this peak quite well by overestimating the peak value by a small amount.
 \begin{figure}[h!]
@@
 
-Figure \ref{fig:velocityU_nocolumn} gives the time history of the streamwise velocity $u$ from geoclaw and experiment. 
-Note that this velocity $u$ in both geoclaw and experiment are depth-averaged. 
-Geoclaw accurately predicted the time of the velocity peak as well as the value of the peak. 
-After the peak arrived, geoclaw overestimated the value of velocity.
-The reason why geoclaw gives a higher velocity may be attributed to its depth-averaged property: water at all depth are propagating with the same speed.
-In the experiment, the velocity gets smaller as it comes closer to the bottom.
+Figure \ref{fig:velocityU_nocolumn} gives the time history of the streamwise velocity $u$ from GeoClaw and experiment. 
+Note that this velocity $u$ in both GeoClaw and experiment are depth-averaged. 
+GeoClaw accurately predicted the time of the velocity peak as well as the value of the peak. 
+After the peak arrived, GeoClaw overestimated the value of velocity.
+The reason why GeoClaw gives a higher velocity may be attributed to its depth-averaged property: Water at all depth are propagating with the same speed.
+In the experiment, the velocity is getting smaller the closer we get to the bottom.
 \begin{figure}[h!]

### 2015-03-10 doc/paper/paper.tex
modified paper.tex

--- 2015-03-09 doc/paper/paper.tex
@@
 \par
-The arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
-If we zoom in \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get \ref{fig:waveheight_nocolumn_zoomin}.
+The arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment and in experiment, the jump is more like a smooth increase. 
+If we zoom in figure \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get \ref{fig:waveheight_nocolumn_zoomin}.
 In \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is about 0.11 second earlier that that from experiment.  
@@
     \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
+    \caption{time history of wave height for case without column}
+    \label{fig:waveheight_nocolumn_zoomin}
+\end{figure}
+\begin{figure}[h!]
+    \centering
     \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
@@
     \label{fig:waveheight_square_x=11.02}
-\end{figure}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
-    \caption{time history of wave height for case with square column, at x = 11.02}
-    \label{fig:waveheight_nocolumn_zoomin}
 \end{figure}

### 2015-03-10 doc/paper/paper.tex

--- 2015-03-09 doc/paper/paper.tex
@@
 If we compare the time interval between first jump and second jump with that in \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
-The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
-This can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
+\par
+The arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment and in experiment, the jump is more like a smooth increase. 
+If we zoom in figure \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get \ref{fig:waveheight_nocolumn_zoomin}.
+In \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is about 0.11 second earlier that that from experiment.  
+The difference in shock shape can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
 In shallow water equations, fluid viscosity is not taken into account, while it's significant in N-S equation (or real physics).
-This causes the shock remain its shape in as it propagates in shallow water equation and smooth out in experiment.
+This causes the shock remain its shape in as it propagates in shallow water equation while smooth out in experiment.
 Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
+\par
 Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
 Thus before the propagating shock arrive, all water downstream the shock stays still. 
-In other words, it even does not know there is a column before the shock hits the column.
-In contrast, N-S equation is a elliptic PDE.
+In other words, it even does not know there is a column before the shock hits the column. In contrast, N-S equation is a elliptic PDE. 
 In real physics govened by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent, which raises the wave hegiht even before the wave hits the column.
@@
     \label{fig:waveheight_square_x=11.0}
+\end{figure}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
+    \caption{time history of wave height for case without column}
+    \label{fig:waveheight_nocolumn_zoomin}
 \end{figure}

### 2015-03-10 dambreak/myplot.py
update myplot.py and setrun.py

@@ -11,8 +11,10 @@ import numpy as np
 def plot_no_column():
     plotdata = ClawPlotData()
     plotdata.outdir = '_output_nocolumn'   # set to the proper output directory
-    gaugeno = 7                   # gauge number to examine
-    g = plotdata.getgauge(gaugeno)
+    gaugeno = 7#this is actually gauge at center of column
+    g1 = plotdata.getgauge(gaugeno)
+    gaugeno = 6#this is actually gauge at farther away from column
+    g2 = plotdata.getgauge(gaugeno)
     #g.t is the array of times
     #g.q is the array of values recorded at the gauges (g.q[m,n] is the m`th variable at time `t[n])
 
@@ -32,7 +34,6 @@ def plot_no_column():
         waveHeightNoColumnOF = list(reader)
         waveHeightNoColumnOF = np.array(waveHeightNoColumnOF)
         waveHeightNoColumnOF = waveHeightNoColumnOF.astype(np.float)
-
     with open('./dataForComp/velocity_nocolumn_exp.csv', 'rb') as csvfile:
         reader = csv.reader(csvfile)
         velocity_nocolumn_exp = list(reader)
@@ -42,30 +43,32 @@ def plot_no_column():
 
     #plot wave height history
     plt.figure()
-    plt.plot(g.t,g.q[0,:],'b-.',label='geoclaw')
+    plt.plot(g1.t,g1.q[0,:],'b-.',label='geoclaw')
     plt.plot(waveHeightNoColumnOF[:,0],waveHeightNoColumnOF[:,1],'r--',label='OpenFOAM')
     plt.plot(waveHeightNoColumnExp[:,0],waveHeightNoColumnExp[:,1],'g--',label='Exp')
     legend = plt.legend(loc='upper center', shadow=True, fontsize='x-large')
+    plt.xlim(3.0,4.0)
     plt.xlabel('time (s)')
     plt.ylabel('wave height (m)')
     plt.title('wave height history at x=11.1')
-    plt.savefig('./myplot/waveheight_nocolumn.png', bbox_inches='tight')
+    #plt.savefig('./myplot/waveheight_nocolumn_xlim_x=11.0.png', bbox_inches='tight')
+    plt.savefig('./myplot/waveheight_nocolumn_zoomin.png', bbox_inches='tight')
     plt.close()
 
-    #plot velocity u history
-    plt.figure()
-    #q[1] is hu instead of u
-    u=np.zeros(g.t.size)
-    for i in range(g.t.size):
[…]

### 2015-03-10 dambreak/setrun.py

@@ -371,10 +371,12 @@ def setgeo(rundata):
     geo_data.coriolis_forcing = False
 
     # == Algorithm and Initial Conditions ==
-    geo_data.sea_level = 0.0
+    geo_data.sea_level = 0.02
+    #geo_data.sea_level = 0.0
     geo_data.dry_tolerance = 1.e-3
     geo_data.friction_forcing = True
-    geo_data.manning_coefficient =  0.015
+    geo_data.manning_coefficient =  0.025
+    #geo_data.manning_coefficient =  0.015
     geo_data.friction_depth = 0.05
 
     # Refinement data
@@ -393,7 +395,7 @@ def setgeo(rundata):
     #   0 => flat bottom
     #   1 => cylinder cloumn
     #   2 => square column
-    topo_choice = 1
+    topo_choice = 0
 
     if (topo_choice == 0):
         topo_column_fname = 'columnZero.tt1'

### 2015-03-10 doc/paper/paper.tex
update paper.tex

--- 2015-03-09 doc/paper/paper.tex
@@
 The shallow water equations are solved using the CLAWPACK software.
+\par
+Three different cases are simulated and compared. In the first case, the column is removed. 
+This is a simple dambreak problem and history of wave height and velocity at where column center should be are measured and compared. 
+In the second case, a square column is set and wave height at some locations ahead the column are measured and compared. 
+For the third case, the square column is replaced with a cylinder with the same data measured and compared. 
+The discrepancy between geoclaw simulation and experiment results are also discussed and analyzed in a deep manner.
 \end{abstract}
@@
 There are two jumps in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
-If we recall that the wave height is around 0.1 in \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak.
+If we recall that the wave height is around 0.1 in figure \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak.
 The second jump is induced by wave reflected by the column.
-If we compare the time interval between first jump and second jump with that in \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
-The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
-This can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
-In shallow water equations, fluid viscosity is not taken into account, while it's significant in N-S equation (or real physics).
-This causes the shock remain its shape in as it propagates in shallow water equation and smooth out in experiment.
+If we compare the time interval between first jump and second jump with that in figure \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
+\par
+There are two distinct differences between geoclaw result and experiment result.
+First, in experiment, the jump is more like a smooth increase.
+Second, arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment.
+\par
+The difference in shock shape can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
+In shallow water equations, fluid viscosity is not taken into account, while it's significant in  real physics.
+This causes the shock remain its shape as it propagates in shallow water equation while smooth out in experiment.
 Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
+\par
+If we zoom in figure \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get figure \ref{fig:waveheight_nocolumn_zoomin}.
+In figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
+This time interval is much smaller than that in figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 
+This difference indicates that the column gives more resistance to shock waves in experiment than it does in shallow water equation model. 
+\par
+A simple analysis to this phenomenon is draw as below.
 Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
@@
 In other words, it even does not know there is a column before the shock hits the column.
+This can be verified by checking history of wave height measured at exactly the same place, 0.04m ahead of leading edge of the square column, but in different cases, one with a square column and one without.
+This comparison is showed in figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
 In contrast, N-S equation is a elliptic PDE. 
-In real physics govened by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent, which raises the wave hegiht even before the wave hits the column.
-This explains why we cannot see two jumps in experiment results.
+In real physics governed by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent. 
+This slows down the propagating speed of shock wave when it comes close to the column in real world and thus shifts the peak toward later time in the plots.
 \begin{figure}[h!]
@@
     \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
+    \caption{time history of wave height for case without column (zoom in)}
+    \label{fig:waveheight_nocolumn_zoomin}
+\end{figure}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_x11_comparison}
+    \caption{comparison of time history of wave height for case with a square column and case without one}
+    \label{fig:waveheight_x11_comparison}
+\end{figure}
+\begin{figure}[h!]
+    \centering
     \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
@@
 \end{figure}
+\par
+Figure \ref{fig:waveheight_square_x=11.04} shows time history of wave height measured at front face of the square column. 
+Difference between arrival time of peak still exists as expected.
+After the peak arrive, the wave gradually becomes stable as indicated by a horizontal line in the plots, in which region the differences between geoclaw and experiment are quite small.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1104_largerTimeRange}
+    \caption{time history of wave height for case with square column, at x = 11.04}
+    \label{fig:waveheight_square_x=11.04}
+\end{figure}
 
 \section{Dam Break with cylinder column}\label{Sec:Cylinder}
-Figure \ref{fig:waveheight_cylinder_x=11.0} is the time history of the wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of the center of the column)
-Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
+Figure \ref{fig:waveheight_cylinder_x=11.0} is time history of the wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of the center of the column)
+%Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
 Thus the result in this case does not agree well with the experiment result.
-Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
+%Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
+Similar feature as what we see in case with a square column can be found in the figure.
+However, the wave height at second peak in this case is almost the same as that in square-column case (around 0.22m in both case), while this feature is quite different for experiment results between two cases.
+\par
+Note that diameter of the cylinder is equal to length of the transection for square column.
+Thus the square column can block the flow in in a more efficient way than cylinder does since cylinder has a smoother shape for the fluid to pass by.
+This explains why we see a lower peak of wave height as well as less steep curve for history of wave height in case with cylinder in experiment. 
+In geoclaw, a geometry is represented with many quadrangular cells.
+In our cylinder case, the cylinder is represented with many square cells since our background mesh is simple structured quadragular mesh with square cells.
+Based on this, we can deduce in geoclaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to flow direction at different locations. 
+Thus it is similar to hitting onto one big face vertical to flow direction, which is the case with square column.  
+This is why my get similar result in geoclaw for case with square column and case with cylinder.
+
 \begin{figure}[h!]
@@
     \label{fig:waveheight_cylinder_x=11.0}
+\end{figure}
+\par
+Figure \ref{fig:waveheight_cylinder_x=11.04} is time history of wave height for cylinder case. As we expected, peak predicted by geoclaw appears earlier but the peak values agree quite well.
+After the peak, geoclaw gives a satisfied prediction of wave height.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x1104_largerTimeScale}
+    \caption{time history of wave height for case with cylinder column, at x = 11.04}
+    \label{fig:waveheight_cylinder_x=11.04}
 \end{figure}

### 2015-03-10 doc/paper/paper.tex
Fixed some typos and style

@@
 In this comparison we will include results from OpenFOAM, an open source CFD software modeling air and water using the Navier-Stokes equations with turbulence model.
-Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height measured at the center location where the column will be mounted. 
-Result from GeoClaw agree quite well in general(Show zoom in plot). At around t=3.8, GeoClaw accurately predicted the arrival of the shock. 
+Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height measured at the center location where the column will be mounted in the following cases. 
+The result from GeoClaw agrees quite well in general(Show zoom in plot). At around t=3.8, GeoClaw accurately predicted the arrival of the shock. 
 GeoClaw overestimates the peak water level a little, while OpenFOAM, gives a better value.
 After a later time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height. 
-GeoClaw also predicts the arrival time of this peak quite well by overestimating the peak value by a small amount.
+GeoClaw also predicts the arrival time for this peak quite well but overestimates the peak value by a small amount.
 \begin{figure}[h!]
@@
 
-Figure \ref{fig:velocityU_nocolumn} gives the time history of the streamwise velocity $u$ from GeoClaw and experiment. 
+Figure \ref{fig:velocityU_nocolumn} shows the time history of the streamwise velocity $u$ from GeoClaw and experiment. 
 Note that this velocity $u$ in both GeoClaw and experiment are depth-averaged. 
 GeoClaw accurately predicted the time of the velocity peak as well as the value of the peak. 
-After the peak arrived, GeoClaw overestimated the value of velocity.
+After the initial jump in the velocity, GeoClaw overestimated the value of velocity.
 The reason why GeoClaw gives a higher velocity may be attributed to its depth-averaged property: Water at all depth are propagating with the same speed.
@@
 \section{Dam Break with square column}\label{Sec:Square}
-Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at 0.04m ahead of the leading edge of the square column.(0.1m ahead of center of the column)
-Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to the previous figures.
-There are two jumps in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
-If we recall that the wave height is around 0.1 in figure \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak.
-The second jump is induced by wave reflected by the column.
-If we compare the time interval between first jump and second jump with that in figure \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
+Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at $0.04m$ ahead of the leading edge of the square column($0.1m$ ahead of center of the column).
+Note that this figure is from $t=3.0$ to $t=4.0$, which shows a shorter range compared to the previous figures.
+There are two jumps in the result given by GeoClaw. A first jump with a change in water level from around $0.02$ to $0.1$ and a second jump from $0.1$ to a higher level of around $0.22$.
+If we recall from Figure \ref{fig:waveheight_nocolumn} that the wave height is around $0.1$, we can confirm that this first jump is caused by the shock wave originate in the dambreak.
+The second jump represents the wave reflected by the column.
+If we compare the time interval from the first to the second jump between measurement location $x=11.0$ in Figure \ref{fig:waveheight_square_x=11.0} and location $11.02$ in Figure \ref{fig:waveheight_square_x=11.02} we can see that as we get closer to the column, this time interval gets shorter.
 \par
-There are two distinct differences between geoclaw result and experiment result.
-First, in experiment, the jump is more like a smooth increase.
-Second, arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment.
+There are two distinct differences between the GeoClaw and the experimental results.
+First, in the experiment, the dambreak causes a smooth increase in water height.
+Second, the arrival time of the peak predicted by GeoClaw is about $0.08$ seconds ahead of the time measured in the experiment.
 \par
-The difference in shock shape can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
-In shallow water equations, fluid viscosity is not taken into account, while it's significant in  real physics.
+The difference in shock shape can be explained by the difference between the shallow water equations that GeoClaw solves and Navier-Stokes equations that better describe the physics.
+In shallow water equations, fluid viscosity is not taken into account, while it is significant in real physics.
 This causes the shock remain its shape as it propagates in shallow water equation while smooth out in experiment.
@@
 \par
-If we zoom in figure \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get figure \ref{fig:waveheight_nocolumn_zoomin}.
-In figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
-This time interval is much smaller than that in figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 
+If we zoom in Figure \ref{fig:waveheight_nocolumn} for smaller time range from $t=3s$ to $t=4s$, we get Figure \ref{fig:waveheight_nocolumn_zoomin}.
+In Figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
+This time interval is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 
 This difference indicates that the column gives more resistance to shock waves in experiment than it does in shallow water equation model. 
@@
 A simple analysis to this phenomenon is draw as below.
-Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
-Thus before the propagating shock arrive, all water downstream the shock stays still. 
+The shallow water equations are essentially a system of hyperbolic PDEs. The solution depends on initial condition along characteristics in the spatio-temporal domain. 
+Thus before the propagating shock arrives, all water downstream the shock has zero velocity. 
 In other words, it even does not know there is a column before the shock hits the column.
-This can be verified by checking history of wave height measured at exactly the same place, 0.04m ahead of leading edge of the square column, but in different cases, one with a square column and one without.
-This comparison is showed in figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
-In contrast, N-S equation is a elliptic PDE. 
-In real physics governed by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent. 
+This can be verified by checking the history of the wave height measured at exactly the same location, $0.04m$ ahead of the leading edge of the square column, but in different cases, one with a square column and one without.
+This comparison is showed in Figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
+In contrast, the Navier-Stokes equations are a system of PDEs of elliptic type. 
+In real physics governed by Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent.
 This slows down the propagating speed of shock wave when it comes close to the column in real world and thus shifts the peak toward later time in the plots.
@@
 \par
-Figure \ref{fig:waveheight_square_x=11.04} shows time history of wave height measured at front face of the square column. 
-Difference between arrival time of peak still exists as expected.
-After the peak arrive, the wave gradually becomes stable as indicated by a horizontal line in the plots, in which region the differences between geoclaw and experiment are quite small.
+Figure \ref{fig:waveheight_square_x=11.04} shows the time history of the wave height measured at the front face of the square column. 
+The difference between the arrival time of the peak still exists as expected.
+After the peak arrives, the wave gradually becomes stable as indicated by a horizontal line in the plot.
+In this region the differences between GeoClaw and experiment are quite small.
 \begin{figure}[h!]
@@
 \section{Dam Break with cylinder column}\label{Sec:Cylinder}
-Figure \ref{fig:waveheight_cylinder_x=11.0} is time history of the wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of the center of the column)
-%Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
+Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the leading edge of cylinder column($0.1m$ ahead of the center of the column).
+%Since the way GeoClaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
 Thus the result in this case does not agree well with the experiment result.
 %Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
-Similar feature as what we see in case with a square column can be found in the figure.
-However, the wave height at second peak in this case is almost the same as that in square-column case (around 0.22m in both case), while this feature is quite different for experiment results between two cases.
+Similar feature as in the square column case Figure \ref{fig:waveheight_square_x=11.0} can be found.
+However, the wave height at the second peak in this case is almost the same as that in square column case (around $0.22m$ in both case), while this feature is quite different for the experimental results between the two cases.
 \par
-Note that diameter of the cylinder is equal to length of the transection for square column.
-Thus the square column can block the flow in in a more efficient way than cylinder does since cylinder has a smoother shape for the fluid to pass by.
-This explains why we see a lower peak of wave height as well as less steep curve for history of wave height in case with cylinder in experiment. 
-In geoclaw, a geometry is represented with many quadrangular cells.
-In our cylinder case, the cylinder is represented with many square cells since our background mesh is simple structured quadragular mesh with square cells.
-Based on this, we can deduce in geoclaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to flow direction at different locations. 
+Note that diameter of the cylinder is equal to half the length of the square column.
+Thus the square column can block the flow in in a more efficient way than cylinder does, since the cylinder has a smoother shape for the fluid to pass by.
+This explains why we see a lower peak of wave height as well as less steep curve for history of wave height in case with the cylinder in the experiment. 
+In GeoClaw, a geometry is represented with quadrilateral cells.
+In the cylinder case, the cylinder is represented with square cells since our background mesh is simple structured quadrilateral mesh with square cells.
+Based on this, we can deduce in GeoClaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to flow direction at different locations. 
 Thus it is similar to hitting onto one big face vertical to flow direction, which is the case with square column.  
-This is why my get similar result in geoclaw for case with square column and case with cylinder.
+This is why my get similar result in GeoClaw for case with square column and case with cylinder.
 
@@
 \par
-Figure \ref{fig:waveheight_cylinder_x=11.04} is time history of wave height for cylinder case. As we expected, peak predicted by geoclaw appears earlier but the peak values agree quite well.
-After the peak, geoclaw gives a satisfied prediction of wave height.
+Figure \ref{fig:waveheight_cylinder_x=11.04} shows the time history of the wave height for the cylinder case.
+As we expect, the peak predicted by GeoClaw appears earlier but the peak values agree quite well with the experiment.
+After the peak, GeoClaw gives a satisfied prediction of the wave height.
 \begin{figure}[h!]

### 2015-03-10 doc/paper/paper.tex
Changes to abstract and title

@@
 
-\title{Comparison of the solution of the shallow equation with experimental}
+\title{Hydraulic bore interaction with a column - A comparison between the solution of the shallow equation and experimental results}
 
@@
 The experiment simulates the interaction between an incident bore and a free-standing coastal structure.
-The shallow water equations are solved using the CLAWPACK software.
+The shallow water equations are solved using the GeoClaw solver of the CLAWPACK software.
 \par
-Three different cases are simulated and compared. In the first case, the column is removed. 
-This is a simple dambreak problem and history of wave height and velocity at where column center should be are measured and compared. 
-In the second case, a square column is set and wave height at some locations ahead the column are measured and compared. 
-For the third case, the square column is replaced with a cylinder with the same data measured and compared. 
-The discrepancy between geoclaw simulation and experiment results are also discussed and analyzed in a deep manner.
+Three different cases are simulated and compared. In the first case a simple dam break problem is performed and evaluated.
+The history of the wave height and velocity at the center location, where the column will be mounted are compared. 
+In the second case, a square column is added and wave height at various locations in front of the column are measured and compared. 
+For the third case, the square column is replaced by a cylinder column and the same variables are measured and compared. 
+The discrepancy between the GeoClaw simulation and the experimental results are discussed and analyzed.
 \end{abstract}

### 2015-03-12 doc/paper/paper.tex
Started to add equation
add todo mark in paper.tex

@@
 \par
+
+%todo
 The difference in shock shape can be explained by the difference between the shallow water equations that GeoClaw solves and Navier-Stokes equations that better describe the physics.
@@
 \par
-If we zoom in Figure \ref{fig:waveheight_nocolumn} for smaller time range from $t=3s$ to $t=4s$, we get Figure \ref{fig:waveheight_nocolumn_zoomin}.
-In Figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
+%todo
+%If we zoom in Figure \ref{fig:waveheight_nocolumn} for smaller time range from $t=3s$ to $t=4s$, we get Figure \ref{fig:waveheight_nocolumn_zoomin}.
+%In Figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
 This time interval is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 

### 2015-03-12 doc/paper/paper.tex
Some more added to the Algo section

--- 2015-03-10 doc/paper/paper.tex
@@
 Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.\\
+\eqm
+\qijnp &=& \qijn - \dtdx \left(\apdqimj^n + \amdqiphj^n\right) - \dtdy \left(\bpdqijm^n + \bmdqijph^n\right)\nono\\
+       &-&\dtdx \left(\tFiphj^n - \tFimhj^n\right) - \dtdy \left(\tGijph^n - \tGijmh^n\right)
+\enm
+In the f-wave approache the flux is decomposed the following way
+\eq
+f(\qi) -f(\qim) = \psum \WZ_{i-1/2}^p = \psum \beta^p_{i-1/2}r^p_{i-1/2}
+\en
+where we write this in one dimensional form since one dimensional Riemann solvers are used.
+Instead of waves $\W$ representing the jump in $Q$ we have now waves $\WZ$ representing jumps in the flux. The fluctuations are defined as
+\eq
+\amdqimh = \sum_{p:s^p_{i-1/2}<0} \WZ_{i-1/2}^p  \quad \text{and} \quad  \apdqimh = \sum_{p:s^p_{i-1/2}>0} \WZ_{i-1/2}^p  
+\en
+for the shallow water equations we are going to have
+\eq
+\begin{bmatrix}
+H_i - H_{i-1}\\
+HU_i - HU_{i-1}\\
+\varphi(\qi) - \varphi(\qim) \\
+B_i - B_{i-1}\\
+\end{bmatrix}
+= \sum_{p=0}^3 \alpha^p_{i-1/2}w^p_{i-1/2}
+\en
+where $\varphi (q) = (hu^2 +\frac{1}{2}gh^2)$ is the momentum flux and $B_i$ and $B_{i-1}$ are cell-averaged approximations of the bathymetry of the respective cell.
+\par
 In the GeoClaw solver adaptive mesh refinement (AMR) can be used.
@@
 \section{Conclusion}\label{Sec:Conclusion}
+
 {\footnotesize

### 2015-03-12 doc/paper/paper.tex
modified paper.tex

--- 2015-03-12 doc/paper/paper.tex
@@
 \usepackage{hyperref}  % for urls and hyperlinks
+%\usepackage[pdftex]{graphicx} 
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.\\
+%todo 
+%there are some compiling problem when I try to compile the following 4 lines
+\eqm
+\qijnp &= \qijn - \dtdx (\apdqimj^n + \amdqiphj^n) - \dtdy (\bpdqimj^n + \bmdqiphj^n)\\
+       &-\dtdx (\tFiphj^n - \tFimhj^n) - \dtdy (\tGijph^n - \tGijmh^n)
+\enm
 In the GeoClaw solver adaptive mesh refinement (AMR) can be used.
@@
 In this comparison we will include results from OpenFOAM, an open source CFD software modeling air and water using the Navier-Stokes equations with turbulence model.
+\par
 Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height measured at the center location where the column will be mounted in the following cases. 
-The result from GeoClaw agrees quite well in general(Show zoom in plot). At around t=3.8, GeoClaw accurately predicted the arrival of the shock. 
+The result from GeoClaw agrees quite well in general. 
+If we zoom in Figure \ref{fig:waveheight_nocolumn} for smaller time range from $t=3s$ to $t=4s$, we get Figure \ref{fig:waveheight_nocolumn_zoomin}.
+In Figure \ref{fig:waveheight_nocolumn_zoomin}, GeoClaw accurately predicted the arrival of the shock at t=3.8s.
+The arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).
 GeoClaw overestimates the peak water level a little, while OpenFOAM, gives a better value.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_nocolumn}
+        \caption{time from 0 to 18}
+        \label{fig:waveheight_nocolumn}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_nocolumn_zoomin}
+        \caption{time from 3 to 4}
+        \label{fig:waveheight_nocolumn_zoomin}
+    \end{subfigure}
     \caption{time history of wave height for case without a column. }
-    \label{fig:waveheight_nocolumn}
 \end{figure}
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.4\textwidth]{./plots/velocityU_nocolumn}
+    \caption{time history of streamwise velocity U for case without a column}
+    \label{fig:velocityU_nocolumn}
+\end{figure}
 Figure \ref{fig:velocityU_nocolumn} shows the time history of the streamwise velocity $u$ from GeoClaw and experiment. 
@@
 In the experiment, the velocity is getting smaller the closer we get to the bottom.
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\textwidth]{./plots/velocityU_nocolumn}
-    \caption{time history of streamwise velocity U for case without a column}
-    \label{fig:velocityU_nocolumn}
-\end{figure}
 
 \section{Dam Break with square column}\label{Sec:Square}
-Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at $0.04m$ ahead of the leading edge of the square column($0.1m$ ahead of center of the column).
-Note that this figure is from $t=3.0$ to $t=4.0$, which shows a shorter range compared to the previous figures.
+%todo
+%add a paragraph giving some introduction here
+Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at $0.04m$ ahead of the front face of the square column($0.1m$ ahead of center of the column).
+%Note that this figure is from $t=3.0$ to $t=4.0$, which shows a shorter range compared to the previous figures.
 There are two jumps in the result given by GeoClaw. A first jump with a change in water level from around $0.02$ to $0.1$ and a second jump from $0.1$ to a higher level of around $0.22$.
@@
 There are two distinct differences between the GeoClaw and the experimental results.
-First, in the experiment, the dambreak causes a smooth increase in water height.
-Second, the arrival time of the peak predicted by GeoClaw is about $0.08$ seconds ahead of the time measured in the experiment.
+(1)In the experiment, the dambreak causes a smooth increase in water height.
+(2)The arrival time of the peak predicted by GeoClaw is about $0.08$ seconds ahead of the time measured in the experiment.
 \par
 
-%todo
-The difference in shock shape can be explained by the difference between the shallow water equations that GeoClaw solves and Navier-Stokes equations that better describe the physics.
+The difference in shock shape can be explained by the difference between the shallow water equations that GeoClaw solves and the real physics of fluid. 
 In shallow water equations, fluid viscosity is not taken into account, while it is significant in real physics.
@@
 \par
-%todo
-%If we zoom in Figure \ref{fig:waveheight_nocolumn} for smaller time range from $t=3s$ to $t=4s$, we get Figure \ref{fig:waveheight_nocolumn_zoomin}.
-%In Figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
 This time interval is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 
@@
 \par
-A simple analysis to this phenomenon is draw as below.
+A simple analysis to this phenomenon is drawn as below.
 The shallow water equations are essentially a system of hyperbolic PDEs. The solution depends on initial condition along characteristics in the spatio-temporal domain. 
@@
 This comparison is showed in Figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
-In contrast, the Navier-Stokes equations are a system of PDEs of elliptic type. 
-In real physics governed by Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent.
+In contrast, the Navier-Stokes equations, which better describes the real physics, are a system of PDEs of elliptic type. 
+In real physics or in Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected (have nonzero velocity) and resist the comming of wave to some extent.
 This slows down the propagating speed of shock wave when it comes close to the column in real world and thus shifts the peak toward later time in the plots.
@@
     \label{fig:waveheight_square_x=11.0}
-\end{figure}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
-    \caption{time history of wave height for case without column (zoom in)}
-    \label{fig:waveheight_nocolumn_zoomin}
 \end{figure}

### 2015-03-12 doc/paper/paper.tex
add plots for waveheight after column and update paper.tex

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x11}
-    \caption{time history of wave height for case with square column}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_square_x11}
+        \caption{at x = 11.0 (0.04m ahead of front face of the square column)}
         \label{fig:waveheight_square_x=11.0}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_square_x1102}
+        \caption{at x = 11.2 (0.02m ahead of front face of the square column)}
+        \label{fig:waveheight_square_x=11.02}
+    \end{subfigure}
+    \caption{time history of wave height for case with a square column.}
 \end{figure}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_x11_comparison}
-    \caption{comparison of time history of wave height for case with a square column and case without one}
-    \label{fig:waveheight_x11_comparison}
-\end{figure}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
-    \caption{time history of wave height for case with square column, at x = 11.02}
-    \label{fig:waveheight_square_x=11.02}
-\end{figure}
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x11}
+%    \caption{time history of wave height for case with square column}
+%    \label{fig:waveheight_square_x=11.0}
+%\end{figure}
 \par
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1104_largerTimeRange}
-    \caption{time history of wave height for case with square column, at x = 11.04}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_x11_comparison}
+        \caption{comparison between case with a square column and case without one}
+        \label{fig:waveheight_x11_comparison}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_square_x1104_largerTimeRange}
+        \caption{case with a square column with a larger time range at x = 11.04}
         \label{fig:waveheight_square_x=11.04}
+    \end{subfigure}
+    \caption{time history of wave height for square column case}
 \end{figure}
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{./plots/waveheight_x11_comparison}
+%    \caption{comparison of time history of wave height for case with a square column and case without one}
+%    \label{fig:waveheight_x11_comparison}
+%\end{figure}
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
+%    \caption{time history of wave height for case with square column, at x = 11.02}
+%    \label{fig:waveheight_square_x=11.02}
+%\end{figure}
+
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1104_largerTimeRange}
+%    \caption{time history of wave height for case with square column, at x = 11.04}
+%    \label{fig:waveheight_square_x=11.04}
+%\end{figure}
 
 \section{Dam Break with cylinder column}\label{Sec:Cylinder}
-Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the leading edge of cylinder column($0.1m$ ahead of the center of the column).
+Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the front point of cylinder column($0.1m$ ahead of the center of the column).
 %Since the way GeoClaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
@@
 Note that diameter of the cylinder is equal to half the length of the square column. 
-Thus the square column can block the flow in in a more efficient way than cylinder does, since the cylinder has a smoother shape for the fluid to pass by.
+Thus in experiment, the square column can block the flow in in a more efficient way than cylinder does, since the cylinder has a smoother shape for the fluid to pass by. 
 This explains why we see a lower peak of wave height as well as less steep curve for history of wave height in case with the cylinder in the experiment. 
@@
 Based on this, we can deduce in GeoClaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to flow direction at different locations. 
+In other words, there is no cell which has a normal vector of front face pointing toward oblique direction other than the direction fluid comes.  
 Thus it is similar to hitting onto one big face vertical to flow direction, which is the case with square column.  
-This is why my get similar result in GeoClaw for case with square column and case with cylinder.
-
+This is why we get similar result in GeoClaw for case with square column and case with cylinder.
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x11}
-    \caption{time history of wave height for case with cylinder column}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x11}
+        \caption{at x = 11}
         \label{fig:waveheight_cylinder_x=11.0}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x1104_largerTimeScale}
+        \caption{at x = 11.04}
+        \label{fig:waveheight_cylinder_x=11.04}
+    \end{subfigure}
+    \caption{time history of wave height for cylinder column case}
 \end{figure}
+
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x11}
+%    \caption{time history of wave height for case with cylinder column}
+%    \label{fig:waveheight_cylinder_x=11.0}
+%\end{figure}
 \par
@@
 After the peak, GeoClaw gives a satisfied prediction of the wave height.
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x1104_largerTimeScale}
-    \caption{time history of wave height for case with cylinder column, at x = 11.04}
-    \label{fig:waveheight_cylinder_x=11.04}
-\end{figure}
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x1104_largerTimeScale}
+%    \caption{time history of wave height for case with cylinder column, at x = 11.04}
+%    \label{fig:waveheight_cylinder_x=11.04}
+%\end{figure}
 

### 2015-03-12 doc/paper/paper.tex

@@
 \eqm
-\qijnp &= \qijn - \dtdx (\apdqimj^n + \amdqiphj^n) - \dtdy (\bpdqimj^n + \bmdqiphj^n)\\
-       &-\dtdx (\tFiphj^n - \tFimhj^n) - \dtdy (\tGijph^n - \tGijmh^n)
+\qijnp &=& \qijn - \dtdx \left(\apdqimj^n + \amdqiphj^n\right) - \dtdy \left(\bpdqijm^n + \bmdqijph^n\right)\nono\\
+       &-&\dtdx \left(\tFiphj^n - \tFimhj^n\right) - \dtdy \left(\tGijph^n - \tGijmh^n\right)
 \enm
+In the f-wave approache the flux is decomposed the following way
+\eq
+f(\qi) -f(\qim) = \psum \WZ_{i-1/2}^p = \psum \beta^p_{i-1/2}r^p_{i-1/2}
+\en
+where we write this in one dimensional form since one dimensional Riemann solvers are used.
+Instead of waves $\W$ representing the jump in $Q$ we have now waves $\WZ$ representing jumps in the flux. The fluctuations are defined as
+\eq
+\amdqimh = \sum_{p:s^p_{i-1/2}<0} \WZ_{i-1/2}^p  \quad \text{and} \quad  \apdqimh = \sum_{p:s^p_{i-1/2}>0} \WZ_{i-1/2}^p  
+\en
+for the shallow water equations we are going to have
+\eq
+\begin{bmatrix}
+H_i - H_{i-1}\\
+HU_i - HU_{i-1}\\
+\varphi(\qi) - \varphi(\qim) \\
+B_i - B_{i-1}\\
+\end{bmatrix}
+= \sum_{p=0}^3 \alpha^p_{i-1/2}w^p_{i-1/2}
+\en
+where $\varphi (q) = (hu^2 +\frac{1}{2}gh^2)$ is the momentum flux and $B_i$ and $B_{i-1}$ are cell-averaged approximations of the bathymetry of the respective cell.
+\par
 In the GeoClaw solver adaptive mesh refinement (AMR) can be used.

### 2015-03-12 doc/paper/paper.tex
update paper.tex. add plots for wave height history measured at back face of column

@@
 Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.\\
-%todo 
-%there are some compiling problem when I try to compile the following 4 lines
 \eqm
@@
 \end{figure}
+\par
+Figure \ref{fig:waveheight_square_back} shows history of wave height measured at back face of the square column (at $X = 11.16$).
+Boundary layer detached near this region thus fluid in this region is difficult to be predicted.
+GeoClaw overestimates wave height in this region.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.4\textwidth]{./plots/waveheight_square_back}
+    \caption{history of wave height measure at back face of the square column}
+    \label{fig:waveheight_square_back}
+\end{figure}
 %\begin{figure}[h!]
@@
 \end{figure}
+\par
+Figure \ref{fig:waveheight_cylinder_back} shows history of wave height measured at back face of the cylinder column (at $X = 11.16$). 
+GeoClaw gives a satisfied prediction in this region.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.4\textwidth]{./plots/waveheight_cylinder_back}
+    \caption{history of wave height measure at back face of the square column}
+    \label{fig:waveheight_cylinder_back}
+\end{figure}
 

### 2015-03-12 doc/paper/paper.tex
Some more changes to the paper and Randys comments

--- 2015-03-12 doc/paper/paper.tex
@@
 In the second case, a square column is added and wave height at various locations in front of the column are measured and compared. 
-For the third case, the square column is replaced by a cylinder column and the same variables are measured and compared. 
+For the third case, the square column is replaced by a cylindrical column and the same variables are measured and compared. 
 The discrepancy between the GeoClaw simulation and the experimental results are discussed and analyzed.
@@
 This is done for the interaction of a bore and a free-standing coastal structure.
-The experiment results are given by Halld\'or \'Arnason \cite{HA}.
+The experimental results are given by Halld\'or \'Arnason \cite{HA}.
 We model the water using the shallow water equations which will be introduced in Section \ref{Sec:ShallowWater}.
@@
 Second we add a square column and compare wave height upstream as well as downstream of the column in \ref{Sec:Square}.
-Lastly in Section \ref{Sec:Cylinder} we compare the wave height at the same location but for a cylinder column.
+Lastly in Section \ref{Sec:Cylinder} we compare the wave height at the same location but for a cylindrical column.
 We close this paper with a conclusion in Section \ref{Sec:Conclusion}.
@@
 This solver combines the qualities of the Roe solver, HLLE-type solver and the f-wave approach.
-From the Roe solver it has the quality of giving a exact solution for a single-shock Riemann problem.
+From the Roe solver it has the quality of giving an exact solution for a single-shock Riemann problem.
 Like the HLLE solver this method is depth positive semidefinite.
 The method has also a natural entropy-fix and gives a better approximation for problems with large rarefactions.
-The solver also maintains large class of steady states.
+The solver also maintains a large class of steady states.
 The topography and drag term are incorporated in the Riemann problem.
 Stability and accuracy are improved by solving a transverse Riemann problem, where the waves moving normal to the cell are split into traverse direction.\\
-%todo 
-%there are some compiling problem when I try to compile the following 4 lines
+In more detail, we have the 2D hyperbolic system
+\eq
+q_t + A(q,x,y)q_x + B(q,x,y)q_y = 0
+\en
+with the wave propagation algorithm
 \eqm
@@
 \enm
-In the f-wave approache the flux is decomposed the following way
+In the f-wave approach the flux is decomposed the following way
 \eq
@@
 \en
-for the shallow water equations we are going to have
+Roe eigenpairs are chosen to include the property of the Roe solver.
+The augmented vector is then split to provide more than $m$ waves since approximating a transonic rarefaction with a single wave can produce non-physical entropy violating numerical solutions.
+For the shallow water equations we have then
 \eq
@@
 In the GeoClaw solver adaptive mesh refinement (AMR) can be used.
-Multiple levels can be use where for each level the domain is specified.
+Multiple levels can be used where for each level the domain is specified.
 The topography for the problem is defined in a file.
 The topography goes into the shallow water equations through the right hand side $B$.
-The value from the file are converted into cell averages using a picewise bilinear function that interpolates the pointwise values.
+The values from the file are converted into cell averages using a picewise bilinear function that interpolates the pointwise values.
 $B$ is then obtained by evaluating the exact integral of this function over the cell.
 \section{The setup of the test case}\label{Sec:Setup}
-The benchmark experiment was conducted in a 16.6 m long, 0.6m wide and 0.45 m deep wave tank. When a gate is lifted, it generates a single bore to impinge the column located downstream.
-We are going to simulate 2 cases in the experiment. The 1st case purely generated the single bore, the 2nd case generated the bore to impinge the column with one side facing the flow.
+The benchmark experiment was conducted in a 16.6 m long, 0.6m wide and 0.45 m deep wave tank. When a gate is lifted, it generates a single bore to impinge on the column located downstream.
+We are going to simulate 2 cases in the experiment.
+In the first case a bore is generated and propagates in the tank without the column.
+In the 2nd case a bore is generated to impinge on the column with one side facing the flow.
 \par
-In the 1st case, wave height and streamwise velocity, U, at different depth, at where the center of column should be located are measured.
-In the 2nd case, the wave height in front and behind the column are measured.
+In the 1st case, wave height and streamwise velocity, U, at different depths, where the center of column should be located are measured.
+In the 2nd case, the wave height in front of and behind the column are measured.
         \begin{figure}[h!]
@@
 
-In all cases below, boundary conditions are all set to reflection wall. 
+In all cases below, boundary conditions are all set to reflecting walls. 
 \section{Dam Break}\label{Sec:Zero}
@@
 The arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).
-GeoClaw overestimates the peak water level a little, while OpenFOAM, gives a better value.
-After a later time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height. 
+GeoClaw overestimates the peak water level a little, while OpenFOAM gives a better value.
+At a later time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height. 
 GeoClaw also predicts the arrival time for this peak quite well but overestimates the peak value by a small amount.
@@
 There are two jumps in the result given by GeoClaw. A first jump with a change in water level from around $0.02$ to $0.1$ and a second jump from $0.1$ to a higher level of around $0.22$.
-If we recall from Figure \ref{fig:waveheight_nocolumn} that the wave height is around $0.1$, we can confirm that this first jump is caused by the shock wave originate in the dambreak.
+If we recall from Figure \ref{fig:waveheight_nocolumn} that the wave height is around $0.1$, we can confirm that this first jump is caused by the shock wave originating from the dambreak.
 The second jump represents the wave reflected by the column.
@@
 Thus before the propagating shock arrives, all water downstream the shock has zero velocity. 
-In other words, it even does not know there is a column before the shock hits the column.
+In other words, it does not know there is a column before the shock hits the column.
 This can be verified by checking the history of the wave height measured at exactly the same location, $0.04m$ ahead of the leading edge of the square column, but in different cases, one with a square column and one without.
 This comparison is showed in Figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
-In contrast, the Navier-Stokes equations, which better describes the real physics, are a system of PDEs of elliptic type. 
+In contrast, the Navier-Stokes equations, which better describes the real physics, are a system of PDEs of parabolic type. 
 In real physics or in Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected (have nonzero velocity) and resist the comming of wave to some extent.
@@
 
-\section{Dam Break with cylinder column}\label{Sec:Cylinder}
-Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the front point of cylinder column($0.1m$ ahead of the center of the column).
+\section{Dam Break with cylindrical column}\label{Sec:Cylinder}
+Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the front point of cylindrical column($0.1m$ ahead of the center of the column).
 %Since the way GeoClaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
@@
 In the cylinder case, the cylinder is represented with square cells since our background mesh is simple structured quadrilateral mesh with square cells.
-Based on this, we can deduce in GeoClaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to flow direction at different locations. 
+Based on this, we can deduce in GeoClaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to the flow direction at different locations. 
 In other words, there is no cell which has a normal vector of front face pointing toward oblique direction other than the direction fluid comes.  
-Thus it is similar to hitting onto one big face vertical to flow direction, which is the case with square column.  
-This is why we get similar result in GeoClaw for case with square column and case with cylinder.
+Thus it is similar to hitting onto one big face vertical to the flow direction, which is the case with a square column.  
+This is why we get similar results in GeoClaw for the case with square column and the case with a cylinder.
 \begin{figure}[h!]
@@
     \end{subfigure}
-    \caption{time history of wave height for cylinder column case}
+    \caption{time history of wave height for cylindrical column case}
 \end{figure}

### 2015-03-13 doc/presentation/pres.tex
Added presentation draft

% Created 2015-03-12 Thu 16:59
\documentclass[xcolor=dvipsnames]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\mode<beamer>{\usetheme{boxes}}
\setbeamertemplate{items}[circle]
\setbeamertemplate{navigation symbols}{}
\usecolortheme[named=RoyalBlue]{structure}
\setbeamercolor{boxcol}{fg=black,bg=black!10}
\setbeamercolor{lowercolgreen}{fg=black,bg=green!10}
\usepackage{pslatex}
\usepackage{amsfonts}
\usepackage{booktabs}
\usepackage{scripttab}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\renewcommand{\vector}     {\boldsymbol}
\newcommand{\uvec}            {\vector{u}}
\newcommand{\vvec}            {\vector{v}}
\newcommand{\wvec}            {\vector{w}}
\newcommand{\svec}            {\vector{s}}
\newcommand{\Bili}[2]      {\left(#1,#2\right)}
\institute{ AM574 Conservation Laws and Finite Volume Methods \\
University of Washington, Seattle USA}
\input{../macros/RJLmacros}
\usetheme{default}
\author{Xinsheng Qin, Kaspar Müller}
\date{March 13, 2015}
\title{Hydraulic bore interaction with a column - A comparison between the solution of the shallow equation and experimental results}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 23.4.1 (Org mode 8.2.5g)}}
\begin{document}

\maketitle
\begin{frame}{Outline}
\tableofcontents
\end{frame}

\usebackgroundtemplate{\includegraphics[width=\paperwidth]{figures/back02.pdf}}

\section{}
\label{sec-1}
\section{Introduction}
\label{sec-2}
\begin{frame}[label=sec-2-1]{Introduction}
\end{frame}

\section{The Model}
\label{sec-3}
\begin{frame}[label=sec-3-1]{The Shallow Water Equations in 2D}
\begin{eqnarray*}
h_t + (uh)_x + (vh)_y &=& 0\nono\\
(hu)_t + (huv)_y +(hu^2 + \frac{1}{2}gh^2)_x &=& -ghB_x - Du\\
(hv)_t + (huv)_x +(hv^2 + \frac{1}{2}gh^2)_y &=& -ghB_y - Dv
\end{eqnarray*}
where $B$ is the topography and $D$ the drag coefficient. $g$ stands for the gravitational acceleration.
\end{frame}

\section{Framework and Method}
\label{sec-4}
\begin{frame}[label=sec-4-1]{CLAWPACK/GeoClaw}
\begin{block}{Software framework}
\end{block}
\begin{block}{Method}
\end{block}
\end{frame}

\section{Test cases}
\label{sec-5}
\begin{frame}[label=sec-5-1]{Setup of the test case}
\begin{columns}
\begin{column}{1.0\textwidth}
\begin{block}{}
\includegraphics[width=.9\linewidth]{../proposal/figures/Diagram_case.jpg} \\
\end{block}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label=sec-5-2]{Case01 - Dam Break}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{block}{\qquad\qquad 0 $\le$ t $\le$ 18}
\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_nocolumn.png} \\
\begin{itemize}
\item Wave arrives at t=3.8s
\item Second jump from reflection from right wall
\end{itemize}
\end{block}
\end{column}

\begin{column}{0.5\textwidth}
\begin{block}{\qquad\qquad 3 $\le$ t $\le$ 4}
\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_nocolumn_zoomin.png}  \\
\begin{itemize}
\item Geoclaw 0.02s ahead
\item Peak value overestimated
\end{itemize}
\vskip5mm
\end{block}
\end{column}
\end{columns}
\end{frame}
\begin{frame}[label=sec-5-3]{Case02 - Dam Break with Square Column}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{block}{}
\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_square_x11.png} \\
\end{block}
\end{column}

\begin{column}{0.5\textwidth}
\begin{block}{}
\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_square_x1102.png}  \\
\end{block}
\end{column}
\end{columns}

\begin{block}{}
\begin{itemize}
\item Second jump reflection from the column
\item Smooth increase in wave height in experiment
\end{itemize}
\end{block}
\end{frame}

\begin{frame}[label=sec-5-4]{Case03 - Dam Break with Cylindrical Column}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{block}{}
\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x11.png} \\
\end{block}
\end{column}

\begin{column}{0.5\textwidth}
\begin{block}{}
\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x1104_largerTimeScale.png}  \\
\vskip5mm
\end{block}
\end{column}
\end{columns}
\end{frame}
\section{Outlook}
\label{sec-6}
\begin{frame}[label=sec-6-1]{Conclusion and Outlook}
\end{frame}
% Emacs 23.4.1 (Org mode 8.2.5g)
\end{document}

### 2015-03-13 doc/paper/paper.tex
Work on paper

@@
 Three different cases are simulated and compared. In the first case a simple dam break problem is performed and evaluated.
-The history of the wave height and velocity at the center location, where the column will be mounted are compared. 
+The history of the wave height and velocity at the center location, where the column will be mounted, are compared. 
 In the second case, a square column is added and wave height at various locations in front of the column are measured and compared. 
@@
 The experimental results are given by Halld\'or \'Arnason \cite{HA}.
-We model the water using the shallow water equations which will be introduced in Section \ref{Sec:ShallowWater}.
-We solve these equations using CLAWPACK's GeoClaw solver which is specialized for geophysical flow problems.
+We model the water using the shallow water equations, which will be introduced in Section \ref{Sec:ShallowWater}.
+We solve these equations using CLAWPACK's GeoClaw solver, which is specialized for geophysical flow problems.
 In Section \ref{Sec:Algo} we show how GeoClaw solves the shallow water equations and how it handles the topography.
@@
 We solve the shallow water equations with CLAWPACK.
-Specifically we used the GeoClaw solver which used a special version of the f-wave approach described in \cite{DLG}.
+Specifically, we used the GeoClaw solver, which used a special version of the f-wave approach described in \cite{DLG}.
 This solver combines the qualities of the Roe solver, HLLE-type solver and the f-wave approach.
@@
 \en
-where we write this in one dimensional form since one dimensional Riemann solvers are used.
+where we write this in one dimensional form, since one dimensional Riemann solvers are used.
 Instead of waves $\W$ representing the jump in $Q$ we have now waves $\WZ$ representing jumps in the flux. The fluctuations are defined as
@@
 Roe eigenpairs are chosen to include the property of the Roe solver.
-The augmented vector is then split to provide more than $m$ waves since approximating a transonic rarefaction with a single wave can produce non-physical entropy violating numerical solutions.
-For the shallow water equations we have then
+The augmented vector is then split to provide more than $m=2$ waves, since approximating a transonic rarefaction with a single wave can produce non-physical entropy violating numerical solutions.
+For the shallow water equations we have then:
 \eq
@@
 where $\varphi (q) = (hu^2 +\frac{1}{2}gh^2)$ is the momentum flux and $B_i$ and $B_{i-1}$ are cell-averaged approximations of the bathymetry of the respective cell.
+The second and third component of $\alpha^p_{i-1/2}w^p_{1-1/2}$ are chosen to ensure conservation.
+These components define $\WZ^p_{i-1/2}$.
+Including the $H_i - H_{i-1}$ splitting allows to create updating fluctuations that corresponds to using a depth positive semidefinite Riemann solver. For the details we refer to \cite{DLG}.
 \par
-In the GeoClaw solver adaptive mesh refinement (AMR) can be used.
-Multiple levels can be used where for each level the domain is specified.
+In the GeoClaw solver, adaptive mesh refinement (AMR) can be used.
+Multiple levels can be used where the domain is specified for each level.
 The topography for the problem is defined in a file.
@@
 \par
-In the 1st case, wave height and streamwise velocity, U, at different depths, where the center of column should be located are measured.
-In the 2nd case, the wave height in front of and behind the column are measured.
+In the 1st case, wave height and streamwise velocity, U, at different depths, where the center of column should be located, are measured.
+In the 2nd case, the wave height in front of and behind the column is measured.
         \begin{figure}[h!]
@@
 
-In all cases below, boundary conditions are all set to reflecting walls. 
+In all cases below, all boundary conditions are set to reflecting walls. 
 \section{Dam Break}\label{Sec:Zero}
@@
 \par
-Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height measured at the center location where the column will be mounted in the following cases. 
+Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height, measured at the center location where the column will be mounted in the following cases. 
 The result from GeoClaw agrees quite well in general. 
@@
 At a later time, the wave gets reflected at the right wall of the tank, which can be seen by the second jump in wave height. 
-GeoClaw also predicts the arrival time for this peak quite well but overestimates the peak value by a small amount.
+GeoClaw also predicts the arrival time for this peak quite well, but overestimates the peak value by a small amount.
 \begin{figure}[h!]
@@
 After the initial jump in the velocity, GeoClaw overestimated the value of velocity.
-The reason why GeoClaw gives a higher velocity may be attributed to its depth-averaged property: Water at all depth are propagating with the same speed.
+The reason why GeoClaw gives a higher velocity may be attributed to its depth-averaged property: water at all depth are propagating with the same speed.
 In the experiment, the velocity is getting smaller the closer we get to the bottom.
@@
 %add a paragraph giving some introduction here
-Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at $0.04m$ ahead of the front face of the square column($0.1m$ ahead of center of the column).
+Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height, measured at $0.04m$ ahead of the front face of the square column($0.1m$ ahead of center of the column).
 %Note that this figure is from $t=3.0$ to $t=4.0$, which shows a shorter range compared to the previous figures.
-There are two jumps in the result given by GeoClaw. A first jump with a change in water level from around $0.02$ to $0.1$ and a second jump from $0.1$ to a higher level of around $0.22$.
+There are two jumps in the result given by GeoClaw: a first jump with a change in water level from around $0.02$ to $0.1$ and a second jump from $0.1$ to a higher level of around $0.22$.
 If we recall from Figure \ref{fig:waveheight_nocolumn} that the wave height is around $0.1$, we can confirm that this first jump is caused by the shock wave originating from the dambreak.
@@
 The difference in shock shape can be explained by the difference between the shallow water equations that GeoClaw solves and the real physics of fluid. 
-In shallow water equations, fluid viscosity is not taken into account, while it is significant in real physics.
-This causes the shock remain its shape as it propagates in shallow water equation while smooth out in experiment.
-Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
+In shallow water equations, fluid viscosity is not taken into account, while in real physics it is significant.
+This causes the shock to retain its shape as it propagates in case of the shallow water equation, while in the experiment the shock smooths out.
+Thus we see a jump in the result from the shallow water equation, instead of a smooth increase as in the experimental result.
 \par
-This time interval is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 
+This time interval is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0}, though data are not measured at exactly the same point for this two plots. 
 This difference indicates that the column gives more resistance to shock waves in experiment than it does in shallow water equation model. 
@@
 The shallow water equations are essentially a system of hyperbolic PDEs. The solution depends on initial condition along characteristics in the spatio-temporal domain. 
-Thus before the propagating shock arrives, all water downstream the shock has zero velocity. 
+Thus, before the propagating shock arrives, all water downstream the shock has zero velocity. 
 In other words, it does not know there is a column before the shock hits the column.
@@
 This comparison is showed in Figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
-In contrast, the Navier-Stokes equations, which better describes the real physics, are a system of PDEs of parabolic type. 
-In real physics or in Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected (have nonzero velocity) and resist the comming of wave to some extent.
-This slows down the propagating speed of shock wave when it comes close to the column in real world and thus shifts the peak toward later time in the plots.
+By contrast, the Navier-Stokes equations, which better describe the real physics, are a system of PDEs of parabolic type. 
+In real physics, or in Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected (have nonzero velocity) and resist the comming of wave to some extent.
+In real world this slows down the propagation speed of the shock wave when it comes close to the column and thus shifts the peak toward later time in the plots.
 \begin{figure}[h!]
@@
 Figure \ref{fig:waveheight_square_back} shows history of wave height measured at back face of the square column (at $X = 11.16$).
-Boundary layer detached near this region thus fluid in this region is difficult to be predicted.
+Boundary layer detached near this region thus fluid in this region is difficult to predict.
 GeoClaw overestimates wave height in this region.
@@
 Thus in experiment, the square column can block the flow in in a more efficient way than cylinder does, since the cylinder has a smoother shape for the fluid to pass by. 
-This explains why we see a lower peak of wave height as well as less steep curve for history of wave height in case with the cylinder in the experiment. 
+This explains why we see a lower peak of wave height, as well as less steep curve for history of wave height in the case with the cylinder in the experiment. 
 In GeoClaw, a geometry is represented with quadrilateral cells. 
 In the cylinder case, the cylinder is represented with square cells since our background mesh is simple structured quadrilateral mesh with square cells.
-Based on this, we can deduce in GeoClaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to the flow direction at different locations. 
-In other words, there is no cell which has a normal vector of front face pointing toward oblique direction other than the direction fluid comes.  
-Thus it is similar to hitting onto one big face vertical to the flow direction, which is the case with a square column.  
+Based on this, we can say that in GeoClaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to the flow direction at different locations. 
+In other words, there is no cell which has a normal vector of front face pointing toward oblique direction other than the direction the fluid comes.  
+Thus, it is similar to hitting onto one big face vertical to the flow direction, which is the case with a square column.  
 This is why we get similar results in GeoClaw for the case with square column and the case with a cylinder.
@@
 Figure \ref{fig:waveheight_cylinder_back} shows history of wave height measured at back face of the cylinder column (at $X = 11.16$). 
-GeoClaw gives a satisfied prediction in this region.
+GeoClaw gives a satisfactory prediction in this region.
 \begin{figure}[h!]
@@
 As we expect, the peak predicted by GeoClaw appears earlier but the peak values agree quite well with the experiment.
-After the peak, GeoClaw gives a satisfied prediction of the wave height.
+After the peak, GeoClaw gives a satisfactory prediction of the wave height.
 %\begin{figure}[h!]
@@
 \section{Conclusion}\label{Sec:Conclusion}
+In this work we compared numerical solutions of the shallow water equations using CLAWPACK's GeoClaw solver with experimental results.
+We saw that GeoClaw underestimates the arrival time of the shock by only 0.02s.
+This is in the same range as OpenFOAM predicts the arrival time.
+The shape of the wave although differs from the experiment and prediction form OpenFOAM.
+Independent of the case the arrival time is predicted always the same in GeoClaw where in the experiment the wave seems to slow down by a small amount if there is an obstacle.
+These discrepancies have to do with the hyperbolic character of the shallow water equations, missing turbulence model as well as depth averaging.
+In general GeoClaw agrees well with the measured water levels at various locations for the square as well as the cylindrical column.
+In the case of the cylindrical column a mapped grid could be used to improve the solution, since in the present case the cylinder shape is approximated by a quadrilateral grid.
+This work could also be extended by comparing the forces on the structure by using $F =\frac{1}{2}C_D\rho d hu^2$.
 

### 2015-03-13 doc/paper/paper.tex
Small fixes in paper.tex

@@
 \par
-Three different cases are simulated and compared. In the first case a simple dam break problem is performed and evaluated.
-The history of the wave height and velocity at the center location, where the column will be mounted, are compared. 
-In the second case, a square column is added and wave height at various locations in front of the column are measured and compared. 
-For the third case, the square column is replaced by a cylindrical column and the same variables are measured and compared. 
-The discrepancy between the GeoClaw simulation and the experimental results are discussed and analyzed.
+We simulate and compare three different cases. In the first case we performe and evaluate a simple dam break problem.
+We compare the history of the wave height and velocity at the center location, where the column will be mounted. 
+In the second case we add a square column and compare wave height at various locations in front of and behind the column. 
+For the third case we replace the square column by a cylindrical column and compare the same variables with the experimental results. 
+We analyze and discuss the discrepancy between the GeoClaw simulation and the experimental results.
 \end{abstract}
@@
 In Figure \ref{fig:waveheight_nocolumn_zoomin}, GeoClaw accurately predicted the arrival of the shock at t=3.8s.
-The arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).
+The arrival time of peak predicted by GeoClaw is almost the same as that from experiment (only about 0.02 second earlier).
 GeoClaw overestimates the peak water level a little, while OpenFOAM gives a better value.

### 2015-03-13 doc/paper/paper.tex
Added Rankine-Hugoniot cond part

@@
 \begin{cases}
-0.25 & x<5.9 \\
+0.25 & x<5.9 \label{eq:init1}\\
 0.02 & \text{otherwise}
 \end{cases}\\
-u_0(x,y) =& 0
+u_0(x,y) =& 0\label{eq:init2}
 \enm
@@
 \par
+If we analyze the Rankine-Hugoniot condition of the initial conditions (\ref{eq:init1}) and (\ref{eq:init2}) for the shallow water equations:
+\eqm
+s(h_*-h) &=& h_*u_* - hu\\
+s(h_*u_*-hu) &=& hu_*^2-hu^2 +\frac{1}{2}g(h_*^2-h^2)
+\enm
+we see that we can not find a $s$ that satisfies both equations, since $u$ and $u*$ are zero.
+This means the Rankine-Hugoniot condition is not satisfied and we expect a right and left going wave originating form the dam break.
 Figure \ref{fig:waveheight_nocolumn} shows the time history of the wave height, measured at the center location where the column will be mounted in the following cases. 

### 2015-03-13 doc/paper/paper.tex
update waveheight plots for cylinder case

--- 2015-03-13 doc/paper/paper.tex
@@
 \par
-This time interval is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0}, though data are not measured at exactly the same point for this two plots. 
+The time interval in no-column case is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0}, though data are not measured at exactly the same point for this two plots. 
 This difference indicates that the column gives more resistance to shock waves in experiment than it does in shallow water equation model. 
@@
     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x11}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x11_finer}
         \caption{at x = 11}

### 2015-03-13 doc/paper/paper.tex
update paper.tex

@@
 Similar feature as in the square column case Figure \ref{fig:waveheight_square_x=11.0} can be found.
-However, the wave height at the second peak in this case is almost the same as that in square column case (around $0.22m$ in both case), while this feature is quite different for the experimental results between the two cases.
+We can see a decrease in peak value for this case compared with the same plots for square column case.
+However, this decrease between two cases from GeoClaw is not as large as that between two cases from experiment.
 \par
@@
 This is why we get similar results in GeoClaw for the case with square column and the case with a cylinder.
+Figure \ref{fig:waveheight_square_x=11.0} also gives results after we refine the mesh. As the mesh becomes finer, the peak value decrease as we expect.
+But with a even finer mesh, we cannot decrease the peak value, which indicates that a mapped grid could be used to improve the solution.
 \begin{figure}[h!]
@@
     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x11_finer}
+        \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x11_finer_new}
         \caption{at x = 11}
@@
 This is in the same range as OpenFOAM predicts the arrival time.
-The shape of the wave although differs from the experiment and prediction form OpenFOAM.
+The shape of the wave although differs from the experiment and prediction from OpenFOAM.
 Independent of the case the arrival time is predicted always the same in GeoClaw where in the experiment the wave seems to slow down by a small amount if there is an obstacle.

### 2015-03-14 doc/presentation/pres.tex
final presentation

@@
-% Created 2015-03-12 Thu 16:59
+% Created 2015-03-13 Fri 15:17
 \documentclass[xcolor=dvipsnames]{beamer}
@@
 \date{March 13, 2015}
-\title{Hydraulic bore interaction with a column - A comparison between the solution of the shallow equation and experimental results}
+\title{Hydraulic bore interaction with a column \\ A comparison between the solution of the shallow equation and experimental results}
 \hypersetup{
@@
 \begin{frame}[label=sec-2-1]{Introduction}
+Comparison between Solution of the shallow water equations solved with CLAWPACK and experimental results.
+\begin{itemize}
+\item Experiment by Halldór Árnason
+\item Comaprison of water level at various locations
+\item 3 Cases
+\begin{enumerate}
+\item Dam Break
+\item Dam Break with square column
+\item Dam Break with cylindrical column
+\end{enumerate}
+\end{itemize}
 \end{frame}
@@
 \label{sec-3}
-\begin{frame}[label=sec-3-1]{The Shallow Water Equations in 2D}
+\begin{frame}[label=sec-3-1]{The Model equations}
+\begin{block}{2D depth averaged  shallow water equations}
 \begin{eqnarray*}
@@
 where $B$ is the topography and $D$ the drag coefficient. $g$ stands for the gravitational acceleration.
+\end{block}
 \end{frame}
@@
 \begin{frame}[label=sec-4-1]{CLAWPACK/GeoClaw}
-\begin{block}{Software framework}
+\begin{block}{GeoClaw}
+\begin{itemize}
+\item Geophysical flow problems
+\item Specialized version of CLAPACK and AMRClaw
+\end{itemize}
 \end{block}
-\begin{block}{Method}
+
+\begin{block}{Requirements for solver}
+\begin{itemize}
+\item Flow over topography
+\item Handle non-trivial steady state
+\begin{itemize}
+\item Ocean at rest
+\end{itemize}
+\item Dry state handling
+\item Multiple scales in space and time
+\end{itemize}
 \end{block}
+\end{frame}
+
+\begin{frame}[label=sec-4-2]{Riemann solver}
+\begin{block}{Approaches}
+\begin{itemize}
+\item f-wave approach
+\begin{itemize}
+\item guarantees numerical conservation
+\item source terms simply included in solver
+\item not clear how to prevent depth negativity
+\end{itemize}
+\item HLLE solver
+\begin{itemize}
+\item guarantees depth non-negativity
+\item fails to capture large transonic rarefactions
+\item not well balanced for steady states
+\end{itemize}
+\end{itemize}
+\end{block}
+
+\begin{block}{Augmented Riemann solver}
+\begin{itemize}
+\item Split into more than two waves
+\begin{itemize}
+\item Add 1 wave to have two waves to represent large transonic rarefactions
+\item Add 1 wave to incorporate topography - balance for steady state
+\end{itemize}
+\end{itemize}
+\end{block}
+\end{frame}
+
+\begin{frame}[label=sec-4-3]{Wave propagation algorithm}
+\begin{itemize}
+\item 2D hyperbolic system
+\end{itemize}
+\begin{equation*}
+q_t + A(q,x,y)q_x + B(q,x,y)q_y = 0
+\end{equation*}
+\begin{itemize}
+\item Approximation of the state
+\end{itemize}
+\begin{eqnarray*}
+\qijnp = \qijn &-& \dtdx \left(\apdqimj^n + \amdqiphj^n\right) \\
+               &-& \dtdy \left(\bpdqijm^n + \bmdqijph^n\right)\\
+               &-&\dtdx \left(\tFiphj^n - \tFimhj^n\right) - \dtdy \left(\tGijph^n - \tGijmh^n\right)
+\end{eqnarray*}
+\begin{itemize}
+\item Fluctuations ${\cal A}^\pm\Delta \Q_{i\mp 1/2,j}^n$ and ${\cal B}^\pm\Delta \Q_{i,j\mp 1/2,j}^n$
+\item Second order correction terms $\tilde F_{i\pm 1/2,j}^n$ and $\tilde G_{i,j\pm 1/2}^n$
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label=sec-4-4]{Approach in 1D}
+\begin{block}{f-wave approach}
+\begin{equation*}
+f(\qi) -f(\qim) = \psum \WZ_{i-1/2}^p = \psum \beta^p_{i-1/2}r^p_{i-1/2}
+\end{equation*}
+\begin{itemize}
+\item jumps in the flux instead of state
+\end{itemize}
+\end{block}
+
+\begin{columns}
+\begin{column}{0.4\textwidth}
+\begin{block}{Fluctuations}
+\begin{equation*}
+\begin{align}
+\amdqimh &= \sum_{p:s^p_{i-1/2}<0} \WZ_{i-1/2}^p \\
+\apdqimh &= \sum_{p:s^p_{i-1/2}>0} \WZ_{i-1/2}^p  
+\end{align}
+\end{equation*}
+\end{block}
+\end{column}
+
+\begin{column}{0.5\textwidth}
+\begin{block}{Decomposition into waves}
+\begin{equation*}
+\begin{bmatrix}
+H_i - H_{i-1}\\
+HU_i - HU_{i-1}\\
+\varphi(\qi) - \varphi(\qim) \\
+B_i - B_{i-1}\\
+\end{bmatrix}
+= \sum_{p=0}^3 \alpha^p_{i-1/2}w^p_{i-1/2}
+\end{equation*}
+\end{block}
+\end{column}
+\end{columns}
 \end{frame}
@@
 
-\begin{column}{0.5\textwidth}
+\begin{column}{0.6\textwidth}
 \begin{block}{\qquad\qquad 3 $\le$ t $\le$ 4}
@@
 \begin{itemize}
-\item Geoclaw 0.02s ahead
+\item GeoClaw 0.02s ahead
 \item Peak value overestimated
@@
 \end{columns}
+
+\begin{block}{}
+\href{../paper/plots/animation/cross_section_nocolumn.mp4}{movie}
+\end{block}
 \end{frame}
+
 \begin{frame}[label=sec-5-3]{Case02 - Dam Break with Square Column}
+\begin{columns}
+\begin{column}{0.5\textwidth}
+\begin{block}{}
+\includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_square_nearcolumn/frame0009fig1.png}\\
+    \includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_square_nearcolumn/frame0010fig1.png}\\
+\end{block}
+\end{column}
+
+\begin{column}{0.5\textwidth}
+\begin{block}{}
+\includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_square_nearcolumn/frame0011fig1.png}\\
+    \includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_square_nearcolumn/frame0012fig1.png}\\
+\end{block}
+\end{column}
+\end{columns}
+
+\begin{block}{}
+\href{../paper/plots/animation/surface_elevation_square_nearcolumn.mp4}{movie}
+\end{block}
+\end{frame}
+
+\begin{frame}[label=sec-5-4]{Case02 - Dam Break with Square Column}
 \begin{columns}
@@
 \begin{itemize}
-\item Second jump reflection from the column
-\item Smooth increase in wave height in experiment
+\item Two jumps in GeoClaw result
+\begin{itemize}
+\item One originates in dam break
+\item Another results from wave reflected by column
+\end{itemize}
+\item Time interval becomes shorter
+\item Difference between GeoClaw result and experiment result:
+\begin{itemize}
+\item Shock shape
+\item Arrival time of the shock
+\end{itemize}
 \end{itemize}
@@
 
-\begin{frame}[label=sec-5-4]{Case03 - Dam Break with Cylindrical Column}
+\begin{frame}[label=sec-5-5]{Case02 - Dam Break with Square Column}
 \begin{columns}
@@
 \begin{block}{}
-\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x11.png} \\
+\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_square_x1104_largerTimeRange.png} \\
+\begin{itemize}
+\item Wave height at front face
+\item Wave height at back face
+\item Comparison between cases with and without column
+\end{itemize}
+\vskip13mm
 \end{block}
@@
 \begin{block}{}
-\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x1104_largerTimeScale.png}  \\
-\vskip5mm
+\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_square_back.png}  \\
+     \includegraphics[width=.9\linewidth]{../paper/plots/waveheight_x11_comparison.png}  \\
 \end{block}
@@
 \end{frame}
+\begin{frame}[label=sec-5-6]{Case03 - Dam Break with Cylindrical Column}
+\begin{columns}
+\begin{column}{0.5\textwidth}
+\begin{block}{}
+\includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_cylinder_nearcolumn/frame0014fig1.png}\\
+    \includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_cylinder_nearcolumn/frame0015fig1.png}\\
+\end{block}
+\end{column}
+
+\begin{column}{0.5\textwidth}
+\begin{block}{}
+\includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_cylinder_nearcolumn/frame0016fig1.png}\\
+    \includegraphics[width=.9\linewidth]{../paper/plots/surface_elevation_cylinder_nearcolumn/frame0017fig1.png}\\
+\end{block}
+\end{column}
+\end{columns}
+
+\begin{block}{}
+\href{../paper/plots/animation/surface_elevation_cylinder_nearcolumn.mp4}{movie}
+\end{block}
+\end{frame}
+
+\begin{frame}[label=sec-5-7]{Case03 - Dam Break with Cylindrical Column}
+\begin{columns}
+\begin{column}{0.5\textwidth}
+\begin{block}{}
+\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x11_finer_new.png} \\
+\begin{itemize}
+\item Mesh refinement
+\item History of wave height at front face and back of the column
+\end{itemize}
+\vskip15mm
+\end{block}
+\end{column}
+
+\begin{column}{0.5\textwidth}
+\begin{block}{}
+\includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x1104_largerTimeScale_finer.png}  \\
+     \includegraphics[width=.9\linewidth]{../paper/plots/waveheight_cylinder_x1116_finer.png} \\
+\end{block}
+\end{column}
+\end{columns}
+
+\begin{block}{}
+\vskip5mm
+\end{block}
+\end{frame}
 \section{Outlook}
@@
 \begin{frame}[label=sec-6-1]{Conclusion and Outlook}
+\begin{itemize}
+\item For the dam break problem, GeoClaw underestimates the arrival time of the shock by only 0.02s
+\item In both cases with cylinder and square column, GeoClaw overestimates the arrival time of the shock
+\item In general GeoClaw agrees well with the measured water levels at various locations for the square as well as the cylindrical column
+\end{itemize}
+\begin{block}{Future work}
+\begin{itemize}
+\item A mapped grid could be used to improve the solution in cylindrical column case
+\item Comparing the forces on the structure by using $F =\frac{1}{2}C_D\rho d hu^2$
+\end{itemize}
+\end{block}
 \end{frame}

### 2015-03-20 doc/paper/paper.tex
correct some grammar error according to comment by Randy

@@
 \par
-We simulate and compare three different cases. In the first case we performe and evaluate a simple dam break problem.
+We simulate and compare three different cases. In the first case we perform and evaluate a simple dam break problem.
 We compare the history of the wave height and velocity at the center location, where the column will be mounted. 
@@
             \includegraphics[width=0.5\textwidth]{../proposal/figures/Diagram_case}
-            \caption{Diagram of experiment apparatus}
+            %\caption{Diagram of experiment apparatus}
+            \caption{Plan view and side view of experiment apparatus}
         \end{figure}
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_nocolumn}
-        \caption{time from 0 to 18}
+        \caption{Time from 0 to 18}
         \label{fig:waveheight_nocolumn}
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_nocolumn_zoomin}
-        \caption{time from 3 to 4}
+        \caption{Time from 3 to 4}
         \label{fig:waveheight_nocolumn_zoomin}
     \end{subfigure}
-    \caption{time history of wave height for case without a column. }
+    \caption{Time history of wave height for case without a column. }
 \end{figure}
@@
     \includegraphics[width=0.4\textwidth]{./plots/velocityU_nocolumn}
-    \caption{time history of streamwise velocity U for case without a column}
+    \caption{Time history of streamwise velocity U for case without a column}
     \label{fig:velocityU_nocolumn}
@@
 The time interval in no-column case is much smaller than that in Figure \ref{fig:waveheight_square_x=11.0}, though data are not measured at exactly the same point for this two plots. 
-This difference indicates that the column gives more resistance to shock waves in experiment than it does in shallow water equation model. 
+This difference indicates that the column gives more resistance to shock waves in the experiment than it does in the shallow water equation model. 
 \par
 A simple analysis to this phenomenon is drawn as below.
-The shallow water equations are essentially a system of hyperbolic PDEs. The solution depends on initial condition along characteristics in the spatio-temporal domain. 
+The shallow water equations are essentially a system of hyperbolic PDEs. The solution depends on initial conditions along characteristics in the spatio-temporal domain. 
 Thus, before the propagating shock arrives, all water downstream the shock has zero velocity. 
@@
 By contrast, the Navier-Stokes equations, which better describe the real physics, are a system of PDEs of parabolic type. 
-In real physics, or in Navier-Stokes equations, when front of the shock wave gets close to the column, water downstream it and upstream column are affected (have nonzero velocity) and resist the comming of wave to some extent.
-In real world this slows down the propagation speed of the shock wave when it comes close to the column and thus shifts the peak toward later time in the plots.
+In real physics, or in Navier-Stokes equations, when the front of the shock wave gets close to the column, water downstream it and upstream column are affected (have nonzero velocity) and resist the coming of wave to some extent.
+In the real world this slows down the propagation speed of the shock wave when it comes close to the column and thus shifts the peak toward later time in the plots.
 \begin{figure}[h!]
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_square_x11}
-        \caption{at x = 11.0 (0.04m ahead of front face of the square column)}
+        \caption{At x = 11.0 (0.04m ahead of front face of the square column)}
         \label{fig:waveheight_square_x=11.0}
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_square_x1102}
-        \caption{at x = 11.2 (0.02m ahead of front face of the square column)}
+        \caption{At x = 11.2 (0.02m ahead of front face of the square column)}
         \label{fig:waveheight_square_x=11.02}
     \end{subfigure}
-    \caption{time history of wave height for case with a square column.}
+    \caption{Time history of wave height for the case with a square column.}
 \end{figure}
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_x11_comparison}
-        \caption{comparison between case with a square column and case without one}
+        \caption{Comparison between case with a square column and case without one}
         \label{fig:waveheight_x11_comparison}
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_square_x1104_largerTimeRange}
-        \caption{case with a square column with a larger time range at x = 11.04}
+        \caption{Case with a square column with a larger time range at x = 11.04}
         \label{fig:waveheight_square_x=11.04}
     \end{subfigure}
-    \caption{time history of wave height for square column case}
+    \caption{Time history of wave height for square column case}
 \end{figure}
@@
 Figure \ref{fig:waveheight_square_back} shows history of wave height measured at back face of the square column (at $X = 11.16$).
-Boundary layer detached near this region thus fluid in this region is difficult to predict.
+The boundary layer is detached near this region and thus fluid in this region is difficult to predict.
 GeoClaw overestimates wave height in this region.
@@
     \includegraphics[width=0.4\textwidth]{./plots/waveheight_square_back}
-    \caption{history of wave height measure at back face of the square column}
+    \caption{History of wave height measured at the back face of the square column}
     \label{fig:waveheight_square_back}
@@
 \section{Dam Break with cylindrical column}\label{Sec:Cylinder}
-Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the front point of cylindrical column($0.1m$ ahead of the center of the column).
+Figure \ref{fig:waveheight_cylinder_x=11.0} shows the time history of the wave height measured at $0.04m$ ahead of the front point of a cylindrical column($0.1m$ ahead of the center of the column).
 %Since the way GeoClaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle.
@@
 %Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
-Similar feature as in the square column case Figure \ref{fig:waveheight_square_x=11.0} can be found.
-We can see a decrease in peak value for this case compared with the same plots for square column case.
+Similar features as in the square column case Figure \ref{fig:waveheight_square_x=11.0} can be found.
+We can see a decrease in peak value for this case compared with the same plots for the square column case.
 However, this decrease between two cases from GeoClaw is not as large as that between two cases from experiment.
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x11_finer_new}
-        \caption{at x = 11}
+        \caption{At x = 11}
         \label{fig:waveheight_cylinder_x=11.0}
@@
         \includegraphics[width=1\textwidth]{./plots/waveheight_cylinder_x1104_largerTimeScale}
-        \caption{at x = 11.04}
+        \caption{At x = 11.04}
         \label{fig:waveheight_cylinder_x=11.04}
     \end{subfigure}
-    \caption{time history of wave height for cylindrical column case}
+    \caption{Time history of wave height for cylindrical column case}
 \end{figure}
@@
     \includegraphics[width=0.4\textwidth]{./plots/waveheight_cylinder_back}
-    \caption{history of wave height measure at back face of the square column}
+    \caption{History of wave height measure at back face of the square column}
     \label{fig:waveheight_cylinder_back}
